### 2023-05-29 doc/paper.tex
Writing the report paper
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

\documentclass[onecolumn,10pt,cleanfoot]{asme2ej}

\usepackage{graphicx} %% for loading jpg figures
\usepackage{bm}
\usepackage{nicefrac}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage{listings}
\usepackage{tablefootnote}
\usepackage{float}
\usepackage{xcolor}
\usepackage{xurl}

\author{Grzegorz Kajda
    \affiliation{
	Bachelor Student, Robotics and \\
	Intelligent Systems\\ \\[-10pt]
	Department of Informatics\\ \\[-10pt]
	The faculty of Mathematics and \\
	Natural Sciences\\ \\[-10pt]
    Email: grzegork@ifi.uio.no
    }
}

\begin{document}

\title{Simulating the Lipkin-Meshov-Glikov model on a classical computer}

\maketitle

\begin{center}
\section{Abstract}

	We simulate the simplified Hamiltonian of the Lipkin-Meshov-Glikov model on a classical computer using the Variational Qunatum Eigensolver (VQE) algorithm. We start by investigating the outcomes resulting from the application of different single qubit quantum gates to qubit systems that are initially prepared in one of the basis states. Subsequently, we construct an entangler circuit and prepare one of the four bell pairs, upon which we apply the Hadamard and CNOT gates. With a fundamental comprehension of the foundational principles governing quantum systems, we contruct quantum circuits for generating ansatz states and meticulously develop the Variational Quantum Eigensolver (VQE) from ground up. We then apply the VQE to two synthetic quantum systems representing respectively one and two particle systems, and compare the performance of the VQE against numerical and analytical solvers. 
\end{center}

\end{document}

### 2023-05-29 doc/paper.tex
Abstract finished
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
     \affiliation{
-	Bachelor Student, Robotics and \\
-	Intelligent Systems\\ \\[-10pt]
-	Department of Informatics\\ \\[-10pt]
-	The faculty of Mathematics and \\
-	Natural Sciences\\ \\[-10pt]
+	Bachelor Student, Robotics and Intelligent Systems\\ \\[-10pt]
+	Department of Informatics The faculty of Mathematics and Natural Sciences\\ \\[-10pt]
     Email: grzegork@ifi.uio.no
@@
 
-	We simulate the simplified Hamiltonian of the Lipkin-Meshov-Glikov model on a classical computer using the Variational Qunatum Eigensolver (VQE) algorithm.
-We start by investigating the outcomes resulting from the application of different single qubit quantum gates to qubit systems that are initially prepared in one of the basis states.
-Subsequently, we construct an entangler circuit and prepare one of the four bell pairs, upon which we apply the Hadamard and CNOT gates.
-With a fundamental comprehension of the foundational principles governing quantum systems, we contruct quantum circuits for generating ansatz states and meticulously develop the Variational Quantum Eigensolver (VQE) from ground up.
-We then apply the VQE to two synthetic quantum systems representing respectively one and two particle systems, and compare the performance of the VQE against numerical and analytical solvers.
+We present a comprehensive simulation of the simplified Hamiltonian of the Lipkin-Meshov-Glikov (LMG) model on a classical computer, employing the Variational Quantum Eigensolver (VQE) algorithm.
+Initially, we explore the outcomes arising from the application of diverse single qubit quantum gates to one qubit systems prepared in basis states.
+Subsequently, an entangler circuit is constructed, and one of the four bell pairs is prepared, to which we apply the Hadamard and CNOT gates.
+Utilizing our newfound grasp of the foundational principles governing quantum systems, we construct quantum circuits for generating ansatz states and methodically develop the VQE from ground up.
+
+To assess the VQE's efficacy, we apply it to synthetic quantum systems representing single and two-particle systems, comparing its performance against numerical and analytical solvers.
+Transitioning to the Lipkin model, we leverage the properties of annihilation and creation operators to reformulate the Lipkin model's Hamiltonian first in terms of the quasi-spin operator and then Pauli operators.
+Employing this encoding scheme, we apply the VQE to systems with spin J=1 and 2, corresponding to Lipkin systems with two and four fermions, respectively.
+We compare the results against exact solutions obtained through standard diagonalization methods.
+This analysis provides valuable insights into the performance and accuracy of the VQE approach in tackling the Lipkin model's Hamiltonian.
+
 \end{center}

### 2023-05-29 doc/paper.tex
Finished abstract and introduction
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 
-\begin{center}
 \section{Abstract}
@@
 
-\end{center}
+\section{Introduction}
+Since its emergence in the early 1900s, quantum mechanics has sparked fervent discussions among physicists, engaging luminaries of modern physics like Einstein, Dirac, Bohr, and many others.
+The discovery of the quantum realm profoundly reshaped our perspective and understanding of the world around us, unveiling a whole new realm of fundamental principles governing the very building blocks of the universe.
+With the advent of computer systems in the latter half of the 20th century, and the introduction of the quantum Turing machine in 1980, the field of quantum computing was born, ultimately leading to the construction of the first quantum computer in 1998.
+Recently, due to technological advances and discovery of new quantum algorithms, we have been granted to an extent, the capability of simulating simple many-body quantum models using classical computers.
+
+An emerging research area in quantum computing focuses on the simulation of nuclear physics, leveraging the unique properties of quantum systems.
+Nuclear systems possess distinct characteristics that make them well-suited for exploration using quantum computers, providing insights into particle interactions.
+However, it is worth noting that quantum computers still face challenges, such as larger error rates compared to classical computers and vulnerability to noise, which can make working with them potentially difficult.
+Although the field of quantum computing is expanding rapidly, widespread accessibility remains limited, with companies like IBM offering only highly restricted access to their quantum technology.
+In light of these factors, we will employ a classical computer to simulate and analyze the simplified Lipkin-Meshkov-Glick model, showcasing the potential that quantum computing holds while acknowledging the current limitations in accessibility.
+
+To facilitate our simulations, we will begin by constructing quantum circuits using single-qubit quantum gates.
+This approach will allow us to implement the Variational Quantum Eigensolver (VQE) algorithm effectively, enabling us to approximate the ground state energy of the Lipkin model for systems composed of two and four particles.
+Apart from the introduction, this paper is organized into four sections.
+In Section Two, we will provide a comprehensive description of the theory and methods employed in this study.
+Following that, we will present our results, and finally, conclude with a summary of our findings.
 

### 2023-05-29 doc/paper.tex
Describing qubits and quantum gates
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 
+ \section{Method}
+
+ \subsection{Quantum gates}
+ The main difference between clasical computers and their quantum counterpart lays in the underlying data they process.
+While a classical computer operates on binary digits (bits), which represent low and high voltage levels, a qunatum computer operates directly on particles such as atoms.
+These leads to different ways of handling the data, as the properties of atoms and the physical theorems that describe their behavior do not allows us to process them in a way identical to how a classical computer processes bits
+
+On a classical machine, data is represented using classical bits and processed using logical gates such as \emph{and, or, xor} and \emph{nand}.
+On a quantum computer, data is represented using \emph{qubits} which are a digital representation of particles.
+The main difference between classical computers and quantum computers lays namely in what data each of these technological wonders operate on.
+T
+
 \end{document}

### 2023-05-29 doc/paper.tex
Intro on qubits finished
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

--- 2023-05-29 doc/paper.tex
@@
 
+ \section{Method}
+
+ \subsection{Quantum gates}
+The primary distinction between classical computers and their quantum counterparts lies in the nature of the data they process.
+In classical computing, the fundamental data unit is the binary digit, commonly known as a bit, which represents different voltage levels associated with low and high states.
+On the other hand, quantum computers utilize particles, such as atoms, as the information unitm leading to a fundamentally different approach to data handling.
+The unique properties of atoms and the underlying principles of quantum mechanics prevent us from treating them in the same manner as classical bits.
+For instance, in classical computing, logical gates like \textit{and}, \textit{or}, \textit{xor}, and \textit{nand} are commonly used for data processing, but these gates produce irreversible outcomes.
+However, on quantum computers, irreversible outcomes are generally not allowed, except in the case of measurement.
+
+In the realm of quantum computing, the fundamental data type is the \textit{qubit}, which represents the digital manifestation of the wave function for a quantum particle.
+The principles of quantum mechanics dictate that the evolution of a quantum system can be mathematically described using unitary operators.
+This implies that the operations applied to quantum systems are inherently reversible.
+Consequently, at any given point in time, a quantum system can be fully characterized by a set of linear unitary operators.
+As a result, data processing on a quantum computer necessitates the use of reversible operators, enabling the "uncomputation" of the outcomes produced by quantum gates.
+The most basic of these operators are typically represented using 2x2 matrices.
+While these 2x2 operators quantum gates operate on individual qubits, by utilizing the Kronecker product (tensor product), we can in theory construct gates for quantum systems composed of arbitrary number of qubits.
+In practice we will often find ourselves restricted to at most 8 or 10 qubits before the computational overhead becomes too high.
+
+The following are the basic quantum gates which we will utilize to contruct more complex gates and ciruits:
+
+\begin{enumerate}
+	\item[\textbf{I}.] \textbf{Identity gate}
+		\begin{equation*}
+			I = \begin{pmatrix}
+			1 & 0 \\
+			0 & 1 \\
+		\end{pmatrix}.
+	\end{equation*}
+
+	\item[\textbf{II.}] \textbf{Pauli-X gate}
+		\begin{equation*}
+			X = \begin{pmatrix}
+			0 & 1 \\
+			1 & 0 \\
+		\end{pmatrix}.
+	\end{equation*}
+
+	\item[\textbf{III.}] \textbf{Pauli-Y gate}
+		\begin{equation*}
+			Y = \begin{pmatrix}
+			0 & -i \\
+			i & 0 \\
+		\end{pmatrix}.
+	\end{equation*}
+
+	\item[\textbf{IV.}] \textbf{Pauli-Z gate}
+		\begin{equation*}
+			Z = \begin{pmatrix}
+			1 & 0 \\
+			0 & -1 \\
+		\end{pmatrix}.
+	\end{equation*}
+
+	\item[\textbf{V.}] \textbf{Hadamard gate}
+		\begin{equation*}
+			H = \frac{1}{\sqrt{2}} \begin{pmatrix}
+			1 & 1 \\
+			1 & -1 \\
+		\end{pmatrix}.
+	\end{equation*}
+
+	\item[\textbf{VI.}] \textbf{Phase gate}
+		\begin{equation*}
+			S = \begin{pmatrix}
+			1 & 0 \\
+			0 & i \\
+		\end{pmatrix}.
+	\end{equation*}
+
+\end{enumerate}
+
+\subsection{Lipkin-Meshkov-Glick model}
+
 \end{document}

### 2023-05-29 doc/paper.tex
Halfway thorugh lipkin model
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 \usepackage{xurl}
+\usepackage{tikz}
+\usepackage{caption}
 
@@
 This approach will allow us to implement the Variational Quantum Eigensolver (VQE) algorithm effectively, enabling us to approximate the ground state energy of the Lipkin model for systems composed of two and four particles.
+
 Apart from the introduction, this paper is organized into four sections.
@@
 
- \section{Method}
+ \section{\textbf{Section II: Theory and Method \\}}
 
@@
 
-In the realm of quantum computing, the fundamental data type is the \textit{qubit}, which represents the digital manifestation of the wave function for a quantum particle.
+On a quantum computer, the fundamental data type used for storing information is the \textit{qubit}, which represents the digital manifestation of the wave function for a quantum particle.
 The principles of quantum mechanics dictate that the evolution of a quantum system can be mathematically described using unitary operators.
@@
 \subsection{Lipkin-Meshkov-Glick model}
+The Lipkin-Meshkov-Glick (LMG) model, often called Lipkin model, is a quantum model describing the collective beahvior of a large number of particles which was introduced in 1964 for the purpose of studying approximation methods used to solve many-body problems.
+Since its introuduction, it has become one of the standard benchmarks [Physical review C] for many-body methods in nuclear and particle physics.
+The model describes a system of N fermions (i.e.
+particles with half-integer spin values like electrons and quarks) which are distributed among two energy levels with quantum numbers $\sigma$ = ±1.
+The upper level has $2\sigma$ = +1 and energy $e_{1}$ = $e/2$, while the lower level has $2\sigma$ = -1 and energy $e_{2}$ = $e/2$.
+This can be visualized in the following way
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+	  % Draw the parallel lines
+	  \draw (0,0) -- (7,0);
+	  \draw (0,1) -- (7,1);
+
+	  % Draw the small circles on the lines
+	  \foreach \x in {2,3}
+	  {
+		\filldraw (\x,1) circle (3pt);
+
+	  }
+	  \foreach \x in {4,5}
+	  {
+		\filldraw (\x,0) circle (3pt);
+	  }
+	\end{tikzpicture}
+	\caption{Figure showing the distrubution of four fermions among two energy levels}
+\end{figure}
+
+The figure depicts two horizontal lines representing the mentioned distinct energy levels, while the dots on the lines represent four fermions that in this case are evenly distributed between these levels.
+Each fermion in the figure corresponds to its own substate, characterized by quantum numbers $p = 1, 2, 3, 4$.
+It is important to note that in the figure, each particle occupies its own unique state.
+This behavior is typical of fermions, which obey the \textit{Pauli exclusion principle} [Basdevant, p.
+348], meaning that two independent fermions cannot occupy the same state at exactly the same time.
+
+Now the Lipkin model is described through its Hamiltonian given by 
+\begin{equation}
+\hat{H} = \hat{H}_{0} + \hat{H}_{1} + \hat{H}_{2}.
+\end{equation}
+
+where each of the the $\hat{H}_{i}$ can be written in terms of the creation and annihilation operators
+
+\begin{equation}
+\hat{H}_0 = \frac{1}{2}\varepsilon\sum_{\sigma,p}\sigma a_{\sigma,p}^{\dagger}a_{\sigma,p},
+\end{equation}
+
+\begin{equation}
+\hat{H}_1 = \frac{1}{2}V\sum_{\sigma,p,p'} a_{\sigma,p}^{\dagger}a_{\sigma,p'}^{\dagger}a_{-\sigma,p'}a_{-\sigma,p},
+\end{equation}
+
+\begin{equation}
+\hat{H}_2 = \frac{1}{2}W\sum_{\sigma,p,p'} a_{\sigma,p}^{\dagger}a_{-\sigma,p'}^{\dagger}a_{\sigma,p'}a_{-\sigma,p}.
+\end{equation}
+
+The role of creation and annihilation operators in a quantum mechanical many-body model is to highten and lower the number of particles in a given substate by one, respectively.
+For instance, applying an creation operator to the upper energy level of \textit{Fig.1} would increase the number of fermions by one in that energy level.
+A useful property of these operators allows us to use them to express the Hamiltonian of the Lipkin model in terms of the so-called \textit{quasi-spin operators} and \textit{number operator}.
+Defining he quasi-spin operators as
+
+\begin{equation}
+\hat{J}_{+} = \sum_{p} a_{p+}^{\dagger}a_{p-},
+\end{equation}
+
+\begin{equation}
+\hat{J}_{-} = \sum_{p} a_{p-}^{\dagger}a_{p+},
+\end{equation}
+
+\begin{equation}
+\hat{J}_{z} = \frac{1}{2}\sum_{p\sigma} \sigma a_{p\sigma}^{\dagger}a_{p\sigma}, 
+\end{equation}
+
+\begin{equation}
+\hat{J}^{2} = J_{+}J_{-} + J_{z}^{2} - J_{z}.
+\end{equation}
 

### 2023-05-29 doc/paper.tex
3/4 of lipkin explanation finished
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 
-The role of creation and annihilation operators in a quantum mechanical many-body model is to highten and lower the number of particles in a given substate by one, respectively.
-For instance, applying an creation operator to the upper energy level of \textit{Fig.1} would increase the number of fermions by one in that energy level.
-A useful property of these operators allows us to use them to express the Hamiltonian of the Lipkin model in terms of the so-called \textit{quasi-spin operators} and \textit{number operator}.
-Defining he quasi-spin operators as
+Where \textit{V} and \textit{W} are constants, $\hat{H}_1$ represents the interaction that can move pairs of fermions between the energy levels, while $\hat{H}_2$ corresponds to the spin-exchange term.
+When it comes to creation and annihilation operators, their role in the context of a quantum mechanical many-body model, is the manipulation of the number of particles in specific substates.
+For instance, applying a creation operator to the upper energy level in Fig.
+1 would increase the number of fermions in that energy level by one.
+One notable property of these operators is their ability to express the Hamiltonian of the Lipkin model in terms of the so-called \textit{quasi-spin operators} and the \textit{number operator}.
+Defining the quasi-spin operators in terms of creation and annihilation operators leads to the following equivalences
 
@@
 
+which when substituted into the equation for the Hamiltonian $\hat{H}$ give us
+
+\begin{equation}
+\hat{H} = \epsilon\hat{J}_z - \frac{1}{2}V(\hat{J}_+\hat{J}_+ + \hat{J}_-\hat{J}_-) + \frac{1}{2}W\left( -N + J_+ J_- + J_- J_+ \right).
+\end{equation}
+
+where the quasi-spin operators obey the commutation relations of angular momentum, thus allwowing us to use them to rewrite the Hamiltonian of the lipkin model (for full proof of the commutators refer to [Fys5429, lecture notes week 8]).
+
 \end{document}

### 2023-05-30 doc/paper.tex
Finishing up the the encoding scheme, TODO: after that, describe diagonalization, power iteration, QR-solver and VQE and move on to presenting the results and conclussion
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 	\end{equation*}
+		For a single qubit, application of each of these matrices represents a rotation of the qubit on the \textit{Bloch sphere (Fig.
+1)}.
+(For more details, take a look at [Hundt, p.56] or [Nielsen, p.19])
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.3\textwidth]{Bloch_sphere.svg.png}
+  \caption{Caption text describing the picture.}
+  \label{fig:example}
+\end{figure}
 
@@
 This behavior is typical of fermions, which obey the \textit{Pauli exclusion principle} [Basdevant, p.
-348], meaning that two independent fermions cannot occupy the same state at exactly the same time.
+348], meaning that two independent fermions cannot occupy the same state at exactly the same time (Do no dwelve on this too much, this is simply the nature of reality sorrounding us).
 
-Now the Lipkin model is described through its Hamiltonian given by 
+Now the Lipkin model is described through its Hamiltonian, a matrix correpsonds to the total energy of a physicial and is fundamental plays a fundamental role in the describing the evolution of a quantum system.
+Here, the Hamiltonian of the system is given by
 \begin{equation}
@@
 
-Where \textit{V} and \textit{W} are constants, $\hat{H}_1$ represents the interaction that can move pairs of fermions between the energy levels, while $\hat{H}_2$ corresponds to the spin-exchange term.
-When it comes to creation and annihilation operators, their role in the context of a quantum mechanical many-body model, is the manipulation of the number of particles in specific substates.
+Where \textit{V} and \textit{W} are constants, $\hat{H}_1$ represents the interaction that can move pairs of fermions between the energy levels, while $\hat{H}_2$ corresponds to the spin-exchange term, scattering one particle up and one down.
+These operators are in turn expressed in term of creation and annihilation operators, whose role in the context of a quantum mechanical many-body model, is the manipulation of the number of particles in specific substates.
 For instance, applying a creation operator to the upper energy level in Fig.
 1 would increase the number of fermions in that energy level by one.
-One notable property of these operators is their ability to express the Hamiltonian of the Lipkin model in terms of the so-called \textit{quasi-spin operators} and the \textit{number operator}.
+One notable property of these operators is their ability to express the Hamiltonian of the Lipkin model in terms of the so-called \textit{quasi-spin operators} and the \textit{number operator}, where the former display behavior analogous to that of spin operators.
 Defining the quasi-spin operators in terms of creation and annihilation operators leads to the following equivalences
@@
 
-which when substituted into the equation for the Hamiltonian $\hat{H}$ give us
+which when substituted into the equation for the Hamiltonian $\hat{H}$ , the equation for \textit{N=2} particles is given by
 
@@
 
-where the quasi-spin operators obey the commutation relations of angular momentum, thus allwowing us to use them to rewrite the Hamiltonian of the lipkin model (for full proof of the commutators refer to [Fys5429, lecture notes week 8]).
+Since the quasi-spin operators exhibit behavior similar to regular spin operators, they follow the commutation relations of angular momentum (for a detailed proof, refer to [Fys5429, lecture notes week 8]).
+This property is of great importance because fermions, such as electrons, are particles with half-integer spin (1/2-spin).
+It means that a fermion possesses intrinsic spin on its own (independent of its motion) [Basdevant, p.275].
+In the context of atomic systems, the energy eigenstates are intimately connected to spin [Nielsen, p.310].
+Therefore, when describing a system composed of atomic particles, it is crucial to consider these internal degrees of freedom.
+
+By utilizing quasi-spin operators, we can account for these internal degrees of freedom and incorporate them into our calculations.
+This allows us to rewrite the Hamiltonian of the Lipkin model in a more convenient form.
+Consequently, we can express the Hamiltonian matrix in terms of the Pauli operators mentioned in the previous subsection.
+This transformation facilitates the analysis and study of the Lipkin model, providing insights into its energy levels and dynamics.
+Before proceeding however, we are going to simplify the Hamiltonian slightly more by setting the spin-exchange term equal to zero, resulting in a dimensionless Hamiltonian
+\begin{equation}
+\hat{H} = \epsilon\hat{J}_z - \frac{1}{2}V(\hat{J}_+\hat{J}_+ + \hat{J}_-\hat{J}_-).
+\end{equation}
+
+Having established the fundamental understanding of our problem and model, we move onto the encoding scheme we will apply to equation (10) for the case of the total spin \textit{J=1,2}, corresponding to situations when the model consists of \textit{N=2,4} particles.
+
+\subsection{Encoding in terms of Pauli matrices}
+As we happend to mention above, the quasi-spin operators in the expression for Hamiltonian of the Lipkin model presented in \textit{equation (10)} can be rewritten in terms of the pauli matrices.
+Using the definitions of the quasi-spin $\hat{J}$ presented in [Basdevant, p.233] and [Fys5419, week8]
+
+\begin{equation}
+\begin{aligned}
+	\hat{J}_x^{(n)} &\rightarrow \frac{X_n}{2}, \\
+	\hat{J}_y^{(n)} &\rightarrow \frac{Y_n}{2}, \\
+	\hat{J}_z^{(n)} &\rightarrow \frac{Z_n}{2}, 
+\end{aligned}
+\end{equation}
+
+where \textit{X, Y} and \textit{Z} are the 2x2 Pauli operators. Having done that, we still need to express the ladder operators in terms of these new mappings
+
+\begin{equation}
+J_{\pm} = J_x \pm iJ_y
+\end{equation}.
+
+With all the definitions, rewritting the Hamiltonian in terms of Pauli matrices becomes a trivial task of substituting in the correct terms, and combining the resulting operators using the a kronecker product.
+The Hamiltonian
+
+\begin{equation}
+H = \frac{1}{2} (Z \otimes I + I \otimes Z) - \frac{v}{2} (X \otimes X - Y \otimes Y)
+\end{equation}
+
+The reason behind our interest in rewritting the Hamiltonian a second time lays in how measurements of the expectation values for particle system are performed on quantum computers.
+Our aim is to measure the state of the particles comprising a quantum system in basis states $|0>$ and $|1>$, which are the eigenstates of the Pauli Z matrix.
 

### 2023-05-30 doc/paper.tex
Finishing up the discussion of the VQE
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 \begin{equation}
-\hat{H} = \epsilon\hat{J}_z - \frac{1}{2}V(\hat{J}_+\hat{J}_+ + \hat{J}_-\hat{J}_-) + \frac{1}{2}W\left( -N + J_+ J_- + J_- J_+ \right).
+	H_I = 2\sum_{p}\hat{J}_z(p) - \frac{V}{2}\sum_{p \neq q}(\hat{J}_+(p)\hat{J}_+(q) + \hat{J}_-(q)\hat{J}_-(p)) - \frac{W}{2}\sum_{p \neq q}(\hat{J}_+(p)\hat{J}_-(q) - \hat{J}_-(p)\hat{J}_+(q))
 \end{equation}
@@
 
-With all the definitions, rewritting the Hamiltonian in terms of Pauli matrices becomes a trivial task of substituting in the correct terms, and combining the resulting operators using the a kronecker product.
-The Hamiltonian
+With all the necessary definitions in place, rewriting the Hamiltonian in terms of Pauli matrices becomes a straightforward task.
+We can substitute the appropriate terms into Equation (10) and combine the resulting operators using the Kronecker product.
+For the Hamiltonian with N=2, the Hamiltonian matrix can be expressed as the sum of the following tensor products:
 
 \begin{equation}
-H = \frac{1}{2} (Z \otimes I + I \otimes Z) - \frac{v}{2} (X \otimes X - Y \otimes Y)
+H = \frac{1}{2} (Z_0 \otimes I + I \otimes Z_1) - \frac{v}{2} (X_1 \otimes X_2 - Y_1 \otimes Y_2), 
 \end{equation}
 
-The reason behind our interest in rewritting the Hamiltonian a second time lays in how measurements of the expectation values for particle system are performed on quantum computers.
-Our aim is to measure the state of the particles comprising a quantum system in basis states $|0>$ and $|1>$, which are the eigenstates of the Pauli Z matrix.
+The reason we are interested in rewriting the Hamiltonian a second time is because of how measurements are performed on quantum computers for particle systems.
+Our goal is to measure the states of the particles in basis states |0⟩ and |1⟩ (or combinations of these states).
+These basis states correspond to the north and south poles of the Bloch sphere (as shown in Figure 1), and interestingly, they are also the eigenstates of the Pauli Z matrix.
+This means that measurements are typically performed along the Z-axis.
+Therefore, in order to perform measurements, we need to express the Lipkin model in terms of the Pauli operators and then rotate the Hamiltonian into the Z-basis using gate equivalences.
+
+\subsection{Analytical method: Diagonalization}
+In solving the Lipkin model, we employ the eigendecomposition technique to find exact solutions by diagonalizing the Hamiltonian matrix.
+This technique relies on the fundamental property that self-adjoint operators, including the Hamiltonian, can be diagonalized.
+This implies the existence of orthogonal eigenvectors denoted by $\mathbf{v}$ and corresponding eigenvalues denoted by $\lambda$, satisfying the equation $\mathbf{A} \mathbf{v} = \lambda \mathbf{v}$.
+
+Since the Hamiltonian is a Hermitian operator [Scherer, p.38], which is self-adjoint, we can apply the eigendecomposition technique to it.
+By solving the characteristic equation $\det(H - \epsilon I) = 0$, where $H$ represents the Hamiltonian matrix and $I$ is the identity matrix, we can determine all the eigenvalues of the Hamiltonian.
+These eigenvalues represent the possible energy states within the Lipkin model.
+
+The eigendecomposition method serves as a powerful tool for obtaining precise solutions to the Lipkin model and gaining insights into its energy spectrum.
+By diagonalizing the Hamiltonian matrix, we can extract valuable information regarding the system's energy levels and associated eigenvectors, which describe the various states of the system.
+
+\subsection{Power iteration}
+In our endeavor to showcase the prowess of the variational quantum algorithm, we will employ the power method for finding an approximate solution to the problem of finding the ground state of the Hamiltonian.
+However, we will tailor it slightly to seek the lowest eigenvalue instead of the largest.
+This method is remarkably straightforward, involving the computation of a matrix-vector product between our target matrix and a randomly initialized vector.
+The resulting product is then assigned to a new vector, which is subsequently normalized.
+This iterative process is repeated for a specified number of iterations or until convergence is attained.
+
+The key insight of the power method is that with each multiplication, the initially initialized vector aligns itself more closely with the eigenvector corresponding to the lowest eigenvalue of the target matrix.
+Convergence is determined by measuring the absolute value of the difference between the product of the current and previous iterations of the algorithm.
+
+By employing the power method with the necessary modifications, we can effectively find the lowest eigenvalue and its corresponding eigenvector, thereby gaining valuable insights into the properties and behavior of the matrix of interest.
+A detailed descripton of the method is presented in [Chapla, p.794].
+
+\subsection{Variational Quantum Eigensolver}
+The Variational Quantum Eigensolver (VQE) is a method used to estimate the ground state energy of a given Hamiltonian, such as the Lipkin Hamiltonian.
+This algorithm is specifically designed to handle Hamiltonian matrices that can be expressed as a sum of a polynomial number of terms involving Pauli operators [Hundt, p.242].
+Examples of such Hamiltonians include the Lipkin model, Ising model, and the Harmonic Oscillator.
+
+To estimate the ground state energy, the VQE algorithm involves three stages.
+In the first stage, an ansatz state is created, which is an initial state parameterized and constructed using single-qubit rotation matrices, namely $R_x$ and $R_y$.
+These rotations correspond to rotations around the x- and y-axes of the Bloch sphere shown in Figure 1.
+If the system consists of multiple qubits, entanglement is introduced among the qubits of the ansatz state using Controlled-NOT gates.
+This is done due to the unknown level of entanglement between the qubits of the system prior to performing measurements.
+
+In the second stage of the algorithm, measurements are performed on the system state according to \textit{Postulate 3} for quantum measurements presented in [Nielsen, p.84-85].
+During measurement, the qubit (or the wave function simulated using a qubit) collapses into one of the possible states of the system.
+This process is repeated a large number of times, and the resulting observations are used to approximate the probabilities associated with each possible substate of the system.
+Since the measurements are performed using a classical computer, it becomes easier to perform a large number of measurements and achieve accurate approximations of the probability amplitudes without the need to physically prepare particles for the system.
+As a result, efficient computation of the probabilities is enabled.
+
+The last and final stage of the VQE finidind optimal paramters for our ansatz, which are the angles $\theta$ and $\phi$ corresponding to the rotations performed along the x- and y axes of the Bloch sphere.
+We could use the naive approach of incrementally iterating going through a large number of angles, and gain a precise approximation to the ground, however this will not suffice for larger many-body systems comprised of dozens of particles.
+Hundt proposes the use of gradient descent ([Hundt, p.
+247]) as means of effectively computing angles for our ansatz.
+Since we
+
+\section{Section III: Results}
 

### 2023-05-30 doc/paper.tex
Discussing results
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 
-The last and final stage of the VQE finidind optimal paramters for our ansatz, which are the angles $\theta$ and $\phi$ corresponding to the rotations performed along the x- and y axes of the Bloch sphere.
-We could use the naive approach of incrementally iterating going through a large number of angles, and gain a precise approximation to the ground, however this will not suffice for larger many-body systems comprised of dozens of particles.
-Hundt proposes the use of gradient descent ([Hundt, p.
-247]) as means of effectively computing angles for our ansatz.
-Since we
+The last and final stage of the VQE algorithm is focused on finding the optimal parameters for our ansatz in order to minimize the expectation value produced by the algorithm.
+The parameters we aim to tune are the angles $\theta$ and $\phi$ corresponding to the rotations around the x- and y-axes of the Bloch sphere.
+A naive approach would involve incrementally iterating through a large number of angles to obtain a precise approximation of the ground state energy over time.
+However, this approach becomes inefficient for larger many-body systems with dozens of particles.
+Hundt proposes the use of gradient descent as a more effective method for approximating the optimal angles for our ansatz [Hundt, p.
+247].
+Fortunately, we can easily define the gradient of our parameterized ansatz.
+The expectation value can be measured in terms of the ansatz as follows:
+
+\begin{equation}
+\langle \psi \vert \hat{H} \vert \psi \rangle = \langle 0 \vert R_i(\phi)^{\dagger} \hat{H} R_i(\phi) \vert 0 \rangle.
+\end{equation}.
+
+If we for instance derivate with respect to the angle $\phi$, we arrive at
+
+\begin{equation}
+\frac{\partial}{\partial \phi} \left[\langle \psi(\theta,\phi) \vert \boldsymbol{H} \vert \psi(\theta,\phi) \rangle\right] = \langle \psi(\theta,\phi) \vert \boldsymbol{H} \left(-\frac{\imath}{2} \boldsymbol{\sigma}_y \vert \psi(\theta,\phi) \rangle + \mathrm{h.c.}\right)
+\end{equation}.
+
+Following the methodology presented in [Fys5419, week11], we arrive at a final expression for the derivative of the expectation value
+
+\begin{equation}
+\langle \psi \vert \boldsymbol{I}^{\dagger}\hat{H}\left(-\frac{\imath}{2}\boldsymbol{\sigma}_i\vert \psi \rangle\right) = \frac{1}{2}\left[\langle \psi \vert R_i\left(\frac{\pi}{2}\right)^{\dagger}\hat{H}R_i\left(\frac{\pi}{2}\right)\vert \psi \rangle - \langle \psi \vert R_i\left(-\frac{\pi}{2}\right)^{\dagger}\hat{H}R_i\left(-\frac{\pi}{2}\right)^{\dagger}\vert \psi \rangle\right] = \frac{1}{2}\left(\langle\hat{H}(\phi+\frac{\pi}{2})\rangle - \langle\hat{H}(\phi-\frac{\pi}{2})\rangle\right).
+\end{equation}.
+
+We will perform measurements and use the following gradient to tune our parameters to arrive at the final value.
 
@@
 
+With the theory and methods explained, we can now present the results achieved by simulating the quanatum gates and variational quantum eigensolver.
+
+\subsection{Application of quantum gates to single qubits basis states}
+
+As the first part of our project, we prepared single qubits in their basis state $|0>$ and $|1>$.
+The effect each gate has on one of the states is presented below, presented in the order the gates were listed in \textbf{Section II}.
+The results of applying the gates are:
+
+		\begin{enumerate}
+			\item[\textbf{I}. Identity gate] 
+				The result of applying both the $\sigma_0$ gate to both $|0>$ and $|1>$ results in the exact same states, as the identity gate is simply the identity matrix.
+Presented using matrices and vectors, the results can be written:
+				\[
+				I \begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				1 & 0 \\
+				0 & 1
+				\end{pmatrix} \\ \\ \\
+				\begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				\]
+
+		\end{enumerate}
+
 \end{document}

### 2023-05-30 doc/paper.tex
Adding the last bit of description about quantum gates
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 \begin{enumerate}
-	\item[\textbf{I}.] \textbf{Identity gate}
+	\item[\textbf{I}.] \textbf{Identity gate} \\
+		This gate does nothing to the state of the qubit, as if a gate never was applied to the qubit in the first place.
 		\begin{equation*}
@@
 
-	\item[\textbf{II.}] \textbf{Pauli-X gate}
+	\item[\textbf{II.}] \textbf{Pauli-X gate} \\
+		The Pauli-X gate, also known as the quantum NOT-gate, flips the probability amplitudes $\alpha$ and $\beta$, meaning that these values swap place in the state vector. 
 		\begin{equation*}
@@
 
-	\item[\textbf{III.}] \textbf{Pauli-Y gate}
+	\item[\textbf{III.}] \textbf{Pauli-Y gate} \\
+		The action of the Pauli Y gate can be described as the mapping of $|0>$ to $i|1>$ and $|1>$ to $i¦0>$.
+The action of the gate can also be viewed as a rotation of the state of the qubit around the y-axis of the Bloch sphere by 180 degrees.
 		\begin{equation*}
@@
 
-	\item[\textbf{IV.}] \textbf{Pauli-Z gate}
+	\item[\textbf{IV.}] \textbf{Pauli-Z gate} \\
+		The Pauli Z gate flips the phase of $|1>$ to $-|1>$ while leaving $|0>$, meaning it flips the relative phase of the qubit. 
 		\begin{equation*}
@@
 
-	\item[\textbf{V.}] \textbf{Hadamard gate}
+	\item[\textbf{V.}] \textbf{Hadamard gate} \\
+		The result of applying the Hadamard gate to either $|0>$ or $|1>$ is a qubit in a super position of both states, with both the basis states having a 50\% of occuring.
+The result is the exactly the same, independent of which of the basis states the qubit is prepared in.
 		\begin{equation*}
@@
 
-	\item[\textbf{VI.}] \textbf{Phase gate}
+	\item[\textbf{VI.}] \textbf{Phase gate} \\
+		The Phase gate, when applied to a qubit, represent the rotation of the second part of a qubit state vector corresponding to $|1>$ by 90 degrees around the z-axis of the bloch sphere.
+
 		\begin{equation*}
@@
 
+	\item[\textbf{VII.}] \textbf{CNOT-gate} \\
 \end{enumerate}
@@
 
-As the first part of our project, we prepared single qubits in their basis state $|0>$ and $|1>$.
-The effect each gate has on one of the states is presented below, presented in the order the gates were listed in \textbf{Section II}.
+As the first step of our project, we prepared single qubits in their basis state $|0>$ and $|1>$ and studied the result the application of the various quantum gates had on these one qubit systems.
+The effect each gate had on the basis states is presented below in the order the gates were listed in \textbf{Section II}.
 The results of applying the gates are:
@@
 		\begin{enumerate}
-			\item[\textbf{I}. Identity gate] 
+			\item[\textbf{I}. ] Identity gate \\
 				The result of applying both the $\sigma_0$ gate to both $|0>$ and $|1>$ results in the exact same states, as the identity gate is simply the identity matrix.
-Presented using matrices and vectors, the results can be written:
+Presented using matrices and vectors, the results can be written.
 				\[
@@
 
+			\item[\textbf{II}. ] Pauli-X gate \\
+				The application of the Pauli X gate result in flipping of the probability amplitudes of the qubit, resulting in $\alpha$ and $\beta$ swapping places in the qubit. 
+				\[
+				I \begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				0 & 1 \\
+				1 & 0
+				\end{pmatrix} \\ \\ \\
+				\begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				\beta \\
+				\alpha
+				\end{pmatrix}
+				\]
+
+				Preparing a qubit in state |0> and applying the quantum NOT-gate results in state |1>.
+
+			\item[\textbf{I}. ] Pauli-X gate \\
+
 		\end{enumerate}

### 2023-05-30 doc/paper.tex
Finishing description for results of applying quanntum gates to single qubits state vectors
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 	\item[\textbf{VII.}] \textbf{CNOT-gate} \\
+		The Controlled-NOT gate, often abbreviated as CNOT, is a more complex gate that combines the Identity and Pauli-X gates using a Kronecker product.
+This gate is designed to operate on quantum systems composed of two qubits rather than a single qubit.
+The name of the gate provides a clue about its functionality.
+By using the first qubit as a control signal, the CNOT gate flips the amplitudes of the second qubit if the control qubit is in the basis state |1>, which corresponds to an active high signal in classical computing.
+However, if the control qubit is in the state |0>, no changes are applied to the second qubit.
+The specific gate we are discussing here is the $CNOT_{0,1}$-gate, where the first qubit serves as the control signal.
+		It's important to note that we can also use the second qubit as the control signal by simply changing the order of the gates in the tensor product used to create the CNOT gate from $I \otimes X$ to $X \otimes I$.
+In matrix form, this operator is represented by a 4x4 matrix that can be written as:
+
+		\begin{equation*}
+		 CNOT_{0,1} = \begin{pmatrix}
+			1 & 0 & 0 & 0 \\
+			0 & 1 & 0 & 0 \\
+			0 & 0 & 0 & 1 \\
+			0 & 0 & 1 & 0 \\
+		\end{pmatrix}
+	\end{equation*}
+
 \end{enumerate}
@@
 				\]
-
 			\item[\textbf{II}. ] Pauli-X gate \\
@@
 				\] 
-
-				Preparing a qubit in state |0> and applying the quantum NOT-gate results in state |1>.
-
-			\item[\textbf{I}. ] Pauli-X gate \\
+			\item[\textbf{III}. ] Pauli-Y gate \\
+				The application of the Y-gate resulted in the exact behavipur described in point III of \textbf{Section II: Quantum gates}, a rotation of the qubits amplitudes bu 180 degrees around the y-axis.
+The result presented as a matrix-vector product is:
+				\[
+				Y \begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				0 & -i \\
+				i & 0
+				\end{pmatrix}
+				\begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				-i\beta \\
+				i\alpha
+				\end{pmatrix}
+				\] 	
+			\item[\textbf{VI}. ] Pauli-Z gate \\
+				Just as expected based on the description provieded in the theory part, the Z-gate flips the relative phase of the state vector of our qubit. The result visualized using matrix-vector product is
+				\[
+				Z \begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				1 & 0 \\
+				0 & -1
+				\end{pmatrix}
+				\begin{pmatrix}
+				\alpha \\
+				\beta
+				\end{pmatrix}
+				=
+				\begin{pmatrix}
+				\alpha \\
+				-\beta
+				\end{pmatrix}
+				\]
+			\item[\textbf{VI}. ] Pauli-Z gate \\
 

### 2023-05-30 doc/paper.tex
Small fix, and finished result description for result of quantum circuits
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 		This gate does nothing to the state of the qubit, as if a gate never was applied to the qubit in the first place.
+Presented below is the matrix representation of the quantum gate the result of applying it to a single qubit:
 		\begin{equation*}
@@
 
+		\begin{equation*}
+		H \begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+		1 & 0 \\
+		0 & 1
+		\end{pmatrix}
+		\begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		\end{equation*}
+
 	\item[\textbf{II.}] \textbf{Pauli-X gate} \\
 		The Pauli-X gate, also known as the quantum NOT-gate, flips the probability amplitudes $\alpha$ and $\beta$, meaning that these values swap place in the state vector.
+Presented below is the matrix representation of the quantum gate the result of applying it to a single qubit:
 		\begin{equation*}
@@
 
+	\begin{equation*}
+		H \begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+		1 & 1 \\
+		1 & -1
+		\end{pmatrix}
+		\begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+		\beta \\
+		\alpha
+		\end{pmatrix}
+	\end{equation*}
+
 	\item[\textbf{III.}] \textbf{Pauli-Y gate} \\
@@
 The action of the gate can also be viewed as a rotation of the state of the qubit around the y-axis of the Bloch sphere by 180 degrees.
+Presented below is the matrix representation of the quantum gate the result of applying it to a single qubit:
+
 		\begin{equation*}
@@
 
+	\begin{equation*}
+		H \begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\frac{1}{\sqrt{2}}\begin{pmatrix}
+		0 & -i \\
+		i & 0
+		\end{pmatrix}
+		\begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\frac{1}{\sqrt{2}}\begin{pmatrix}
+		-i\beta \\
+		i\alpha
+		\end{pmatrix}
+	\end{equation*}
+
 	\item[\textbf{IV.}] \textbf{Pauli-Z gate} \\
 		The Pauli Z gate flips the phase of $|1>$ to $-|1>$ while leaving $|0>$, meaning it flips the relative phase of the qubit.
+Presented below is the matrix representation of the quantum gate the result of applying it to a single qubit:
+
 		\begin{equation*}
@@
 		\end{pmatrix}.
+	\end{equation*}
+
+		\begin{equation*}
+			Z \begin{pmatrix}
+			\alpha \\
+			\beta
+			\end{pmatrix}
+			=
+			\begin{pmatrix}
+			1 & 0 \\
+			0 & -1
+			\end{pmatrix}
+			\begin{pmatrix}
+			\alpha \\
+			\beta
+			\end{pmatrix}
+			=
+			\begin{pmatrix}
+			\alpha \\
+			-\beta
+			\end{pmatrix}
 	\end{equation*}
@@
 
+		\begin{equation*}
+			H \begin{pmatrix}
+			\alpha \\
+			\beta
+			\end{pmatrix}
+			=
+			\frac{1}{\sqrt{2}}\begin{pmatrix}
+			1 & 1 \\
+			1 & -1
+			\end{pmatrix}
+			\begin{pmatrix}
+			\alpha \\
+			\beta
+			\end{pmatrix}
+			=
+			\frac{1}{\sqrt{2}}\begin{pmatrix}
+			\alpha + \beta \\
+			\alpha - \beta
+			\end{pmatrix}
+	\end{equation*}
+
 	\item[\textbf{VI.}] \textbf{Phase gate} \\
@@
 	\end{equation*}
-		For a single qubit, application of each of these matrices represents a rotation of the qubit on the \textit{Bloch sphere (Fig.
-1)}.
-(For more details, take a look at [Hundt, p.56] or [Nielsen, p.19])
 
-\begin{figure}[ht]
+	\begin{equation*}
+		S \begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+		1 & 0 \\
+		0 & i
+		\end{pmatrix}
+		\begin{pmatrix}
+		\alpha \\
+		\beta
+		\end{pmatrix}
+		=
+		\begin{pmatrix}
+		\alpha \\
+		i\beta
+		\end{pmatrix}
+	\end{equation*}
+
+	For a single qubit, application of each of these matrices represents a rotation of the qubit on the \textit{Bloch sphere (Fig. 1)}. (For more details, take a look at [Hundt, p.56] or [Nielsen, p.19])
+
+\begin{figure}[H]
   \centering
   \includegraphics[width=0.3\textwidth]{Bloch_sphere.svg.png}
-  \caption{Caption text describing the picture.}
+  \caption{Three dimensional sphere, known as the Bloch sphere, used to visualize the state of a qubit and the action each quantum gate has on the state vector of the qubit.}
   \label{fig:example}
@@
 \subsection{Application of quantum gates to single qubits basis states}
+In the initial phase of our project, we embarked on the exploration of single qubits by preparing them in their respective basis states $|0\rangle$ and $|1\rangle$.
+Our objective was to investigate the effects of different quantum gates on these individual qubit systems.
+Notably, the outcomes observed align precisely with the theoretical predictions discussed in the earlier sections of this report.
+Therefore, we present the resulting states obtained by applying these gates to a qubit prepared in the state $|0\rangle$ in the table below, which serves as a comprehensive overview of our experimental findings.
 
-As the first step of our project, we prepared single qubits in their basis state $|0>$ and $|1>$ and studied the result the application of the various quantum gates had on these one qubit systems.
-The effect each gate had on the basis states is presented below in the order the gates were listed in \textbf{Section II}.
-The results of applying the gates are:
+\begin{center}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{|c|c|c|}
+\hline
+Initial State & Quantum Gate & Final State \\
+\hline
+$\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ & $I = \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}$ & $\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ \\
+\hline
+$\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ & $X = \begin{pmatrix} 0 & 1 \\ 1 & 0 \end{pmatrix}$ & $\begin{pmatrix} 0 \\ 1 \end{pmatrix}$ \\
+\hline
+$\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ & $Y = \begin{pmatrix} 0 & -i \\ i & 0 \end{pmatrix}$ & $i\begin{pmatrix} 0 \\ 1 \end{pmatrix}$ \\
+\hline
+$\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ & $Z = \begin{pmatrix} 1 & 0 \\ 0 & -1 \end{pmatrix}$ & $\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ \\
+\hline
+$\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ & $H = \frac{1}{\sqrt{2}}\begin{pmatrix} 1 & 1 \\ 1 & -1 \end{pmatrix}$ & $\frac{1}{\sqrt{2}}\begin{pmatrix} 1 \\ 1 \end{pmatrix}$ \\
+\hline
+$\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ & $S = \begin{pmatrix} 1 & 0 \\ 0 & i \end{pmatrix}$ & $\begin{pmatrix} 1 \\ 0 \end{pmatrix}$ \\
+\hline
+\end{tabular}
+\captionof{table}{Results of applying listed gates to a qubit prepared in the basis state $|0\rangle$.}
+\end{center}
 
-		\begin{enumerate}
-			\item[\textbf{I}. ] Identity gate \\
-				The result of applying both the $\sigma_0$ gate to both $|0>$ and $|1>$ results in the exact same states, as the identity gate is simply the identity matrix.
-Presented using matrices and vectors, the results can be written.
-				\[
-				I \begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				1 & 0 \\
-				0 & 1
-				\end{pmatrix} \\ \\ \\
-				\begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				\]
-			\item[\textbf{II}. ] Pauli-X gate \\
-				The application of the Pauli X gate result in flipping of the probability amplitudes of the qubit, resulting in $\alpha$ and $\beta$ swapping places in the qubit. 
-				\[
-				I \begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				0 & 1 \\
-				1 & 0
-				\end{pmatrix} \\ \\ \\
-				\begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				\beta \\
-				\alpha
-				\end{pmatrix}
-				\] 
-			\item[\textbf{III}. ] Pauli-Y gate \\
-				The application of the Y-gate resulted in the exact behavipur described in point III of \textbf{Section II: Quantum gates}, a rotation of the qubits amplitudes bu 180 degrees around the y-axis.
-The result presented as a matrix-vector product is:
-				\[
-				Y \begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				0 & -i \\
-				i & 0
-				\end{pmatrix}
-				\begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				-i\beta \\
-				i\alpha
-				\end{pmatrix}
-				\] 	
-			\item[\textbf{VI}. ] Pauli-Z gate \\
-				Just as expected based on the description provieded in the theory part, the Z-gate flips the relative phase of the state vector of our qubit. The result visualized using matrix-vector product is
-				\[
-				Z \begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				1 & 0 \\
-				0 & -1
-				\end{pmatrix}
-				\begin{pmatrix}
-				\alpha \\
-				\beta
-				\end{pmatrix}
-				=
-				\begin{pmatrix}
-				\alpha \\
-				-\beta
-				\end{pmatrix}
-				\]
-			\item[\textbf{VI}. ] Pauli-Z gate \\
+As part of our expermients with quantum circuits, we also constructed the entangler circuit and prepared the first bell state $\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$.
+Thereafter we applied the Hadamard gate to first and second gates separetly, before proceeding with the application of CNOT(0,1)-gate, and ending with the application of the Hadamard and CNOT(0,1) gates in the reverse order to that of the entangler circuit.
+The last operation resulted in the uncomputation of the entanglement.
+All results are presented in the table below.
 
-		\end{enumerate}
+\begin{center}
+\begin{tabular}{|c|c|c|}
+\hline
+First Bell Pair & Quantum Operators & Resulting State \\
+\hline
+$\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$ & $H \otimes I$ & $\frac{1}{\sqrt{2}}(|00\rangle + |10\rangle + |01\rangle + |11\rangle)$ \\
+$\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$ & $I \otimes H$ & $\frac{1}{\sqrt{2}}(|00\rangle + |10\rangle + |01\rangle + |11\rangle)$ \\
+$\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$ & CNOT(0,1) & $\frac{1}{\sqrt{2}}(|00\rangle + |01\rangle)$ \\
+$\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$ & $H \otimes$ CNOT(0,1) & $|00\rangle$ \\
+$\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$ & CNOT(0,1)adj$\otimes$Hadj & $|00\rangle$ \\
+\hline
+\end{tabular}
+\end{center}
+
+\subsection{Solving the ine qubit synthetic Hamiltonian}
 

### 2023-05-30 src/runs/bell_task_a.py

@@ -155,7 +155,9 @@ if __name__ == "__main__":
     H_b00 = q0_Hadamard @ b00
     print(H_b00, "\n")
 
-    print("Measurements after application of CNOT(0, 1)-gate")
+    print(
+        "Measurements after application of Hadamard-gate to the first qubit of the bell pair"
+    )
     print("-------------------------------------------------------------\n")
     psi, outcome, counts, meas_probs = measure(H_b00, 1000)
     print("State of the system after measurement: ")
@@ -222,7 +224,7 @@ if __name__ == "__main__":
     print("Result of applying the adjoint of the gates to the b00 state: ")
     print(Hadamard_adj @ CNOT_adj @ b00, "\n")
 
-    print("Measurements after application of CNOT(0, 1) and Hadamar gate to b00")
+    print("Measurements after application of CNOT(0, 1) and Hadamard gate to b00")
     print("-------------------------------------------------------------\n")
     psi, outcome, counts, meas_probs = measure(HCNOT_b00, 1000)
     print("State of the system after measurement: ")

### 2023-05-31 doc/paper.tex
describing the one body Hamiltonian
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 
-To assess the VQE's efficacy, we apply it to synthetic quantum systems representing single and two-particle systems, comparing its performance against numerical and analytical solvers.
+To assess the VQE's efficacy, we apply it to quantum systems representing single and two-particle systems, comparing its performance against numerical and analytical solvers.
 Transitioning to the Lipkin model, we leverage the properties of annihilation and creation operators to reformulate the Lipkin model's Hamiltonian first in terms of the quasi-spin operator and then Pauli operators.
@@
 
+\subsection{One and two body Hamiltonians} 
+Prior to the simulation of the Lipkin model, we are going to simulate two simple Hamiltonians using our VQE algoirthm, and access its performance by comparing the achieved results against the eigenvalues found using standard eigenvalue decomposition.
+The basic systems we consider are described below.
+
+\subsubsection{One body Hamiltonian}
+We define a symmetric matrix $H \in \mathbb{R}^{2 \times 2}$ as:
+\[
+H =
+\begin{bmatrix}
+H_{11} & H_{12} \\
+H_{21} & H_{22}
+\end{bmatrix},
+\]
+We let $H = H_0 + H_I$, where:
+\[
+H_0 =
+\begin{bmatrix}
+E_1 & 0 \\
+0 & E_2
+\end{bmatrix},
+\]
+is a diagonal matrix. Similarly,
+\[
+H_I =
+\begin{bmatrix}
+V_{11} & V_{12} \\
+V_{21} & V_{22}
+\end{bmatrix},
+\]
+where $V_{ij}$ represent various interaction matrix elements. We can view $H_0$ as the non-interacting solution:
+\[
+H_0|0\rangle = E_1|0\rangle, \quad (1)
+\]
+and
+\[
+H_0|1\rangle = E_2|1\rangle, \quad (2)
+\]
+where we have defined the orthogonal computational one-qubit basis states $|0\rangle$ and $|1\rangle$.
+
+We rewrite $H$ (and $H_0$ and $H_I$) via Pauli matrices:
+\[
+H_0 = E I + \Omega \sigma_z, \quad E = \frac{E_1 + E_2}{2}, \quad \Omega = \frac{E_1 - E_2}{2},
+\]
+and
+\[
+H_I = cI + \omega_z \sigma_z + \omega_x \sigma_x,
+\]
+with $c = \frac{V_{11} + V_{22}}{2}$, $\omega_z = \frac{V_{11} - V_{22}}{2}$, and $\omega_x = V_{12} = V_{21}$. We let our Hamiltonian depend linearly on a strength parameter $\lambda$:
+\[
+H = H_0 + \lambda H_I,
+\]
+with $\lambda \in [0, 1]$, where the limits $\lambda = 0$ and $\lambda = 1$ represent the non-interacting (or unperturbed) and fully interacting system, respectively.
+The model is an eigenvalue problem with only two available states.
+
+Here we set the parameters $E_1 = 0$, $E_2 = 4$, $V_{11} = -V_{22} = 3$, and $V_{12} = V_{21} = 0.2$.
+The non-interacting solutions represent our computational basis.
+Pertinent to our choice of parameters, at $\lambda \geq \frac{2}{3}$, the lowest eigenstate is dominated by $|1\rangle$ while the upper is $|0\rangle$.
+At $\lambda = 1$, the $|0\rangle$ mixing of the lowest eigenvalue is 1\%, while for $\lambda \leq \frac{2}{3}$, we have a $|0\rangle$ component of more than 90\%.
+The character of the eigenvectors has therefore been interchanged when passing $\lambda = \frac{2}{3}$.
+The value of the parameter $V_{12}$ represents the strength of the coupling between the two states.
+
+\subsubsection{Two-body Hamiltonian}
+
 \subsection{Lipkin-Meshkov-Glick model}
@@
 
-As part of our expermients with quantum circuits, we also constructed the entangler circuit and prepared the first bell state $\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$.
-Thereafter we applied the Hadamard gate to first and second gates separetly, before proceeding with the application of CNOT(0,1)-gate, and ending with the application of the Hadamard and CNOT(0,1) gates in the reverse order to that of the entangler circuit.
-The last operation resulted in the uncomputation of the entanglement.
-All results are presented in the table below.
+During our exploration of quantum circuits, we akso conducted experiments involving the construction of an entangler circuit and the preparation of the first Bell state, $\frac{1}{\sqrt{2}}(|00\rangle + |11\rangle)$.
+Subsequently, we applied the Hadamard gate to the first and second qubits individually.
+Following this, we proceeded with the application of the CNOT(0,1) gate and concluded by reversing the order of the Hadamard and CNOT(0,1) gates, in contrast to the entangler circuit.
+Notably, this final operation resulted in the uncomputation of the entanglement.
+The comprehensive results of these operations are presented in the table below.
 
@@
 
-\subsection{Solving the ine qubit synthetic Hamiltonian}
+\subsection{Solving one and two qubit Hamiltonians} 
+Prior to simulating the Lipkin model using our VQE alogorithm, we constructed two simple Hamiltonians each representing a system comprised of one and two qubits respectively.
+We started by setting up the one qubit system where the Hamiltonian matrix is given by the simple eqution constituting a non-interacting part $H_0$ and an interacting part $H_1$,
 

### 2023-05-31 doc/paper.tex
Added subsections on one- and two-body Hamiltonians and mentioned the von neumann entropy for the two-body Hamiltonian in the theory part of the paper
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 The value of the parameter $V_{12}$ represents the strength of the coupling between the two states.
+([Fys5419, project 1])
 
 \subsubsection{Two-body Hamiltonian}
+We also introduce a simple system composed of two qubits, where the system can be thought of as composed of two subsystems A and B.
+Each subsystem has computational basis states
+\begin{equation}
+|0\rangle_{A,B} = \begin{bmatrix} 1 \\ 0 \end{bmatrix},
+\end{equation}
+\begin{equation}
+|1\rangle_{A,B} = \begin{bmatrix} 0 \\ 1 \end{bmatrix}.
+\end{equation}
+The subsystems could represent single particles or composite many-particle
+systems of a given symmetry. This leads to the many-body computational basis
+states
+\begin{equation}
+|00\rangle = |0\rangle_A \otimes |0\rangle_B = \begin{bmatrix} 1 \\ 0 \\ 0 \\ 0 \end{bmatrix},
+\end{equation}
+\begin{equation}
+|01\rangle = |0\rangle_A \otimes |1\rangle_B = \begin{bmatrix} 0 \\ 1 \\ 0 \\ 0 \end{bmatrix},
+\end{equation}
+\begin{equation}
+|10\rangle = |1\rangle_A \otimes |0\rangle_B = \begin{bmatrix} 0 \\ 0 \\ 1 \\ 0 \end{bmatrix},
+\end{equation}
+and finally
+\begin{equation}
+|11\rangle = |1\rangle_A \otimes |1\rangle_B = \begin{bmatrix} 0 \\ 0 \\ 0 \\ 1 \end{bmatrix}.
+\end{equation}
+These computational basis states also define the eigenstates of the non-interacting Hamiltonian
+\begin{equation}
+H_0 |00\rangle = \epsilon_{00} |00\rangle,
+\end{equation}
+\begin{equation}
+H_0 |10\rangle = \epsilon_{10} |10\rangle,
+\end{equation}
+\begin{equation}
+H_0 |01\rangle = \epsilon_{01} |01\rangle,
+\end{equation}
+and
+\begin{equation}
+H_0 |11\rangle = \epsilon_{11} |11\rangle.
+\end{equation}
+The interacting part of the Hamiltonian \(H_I\) is given by the tensor product of
+two \(\sigma_x\) and \(\sigma_z\) matrices, respectively, that is
+\begin{equation}
+H_I = H_x \sigma_x \otimes \sigma_x + H_z \sigma_z \otimes \sigma_z,
+\end{equation}
+where \(H_x\) and \(H_z\) are interaction strength parameters. Our final Hamiltonian
+matrix is given by
+\begin{equation}
+H = \begin{bmatrix}
+\epsilon_{00} + H_z & 0 & 0 & H_x \\
+0 & \epsilon_{10} - H_z & H_x & 0 \\
+0 & H_x & \epsilon_{01} - H_z & 0 \\
+H_x & 0 & 0 & \epsilon_{11} + H_z
+\end{bmatrix}.
+\end{equation}
+
+\subsubsection{Von Neumann entropy for the two-body Hamiltonian}
+The four eigenstates of the above Hamiltonian matrix can in turn be used to define density matrices.
+As an example, the density matrix of the first eigenstate (lowest energy \(E_0\), denoted as \(\Psi_0\)) is
+
+\begin{equation}
+\rho_0 = (\alpha_{00}|00\rangle\langle00| + \alpha_{10}|10\rangle\langle10| + \alpha_{01}|01\rangle\langle01| + \alpha_{11}|11\rangle\langle11|)
+\end{equation}
+
+where the coefficients \(\alpha_{ij}\) are the eigenvector coefficients resulting from the solution of the above eigenvalue problem.
+
+We can then in turn define the density matrix for the subsets A or B as
+
+\begin{equation}
+\rho_A = \text{Tr}_B(\rho_0) = \langle0|\rho_0|0\rangle_B + \langle1|\rho_0|1\rangle_B,
+\end{equation}
+
+or
+
+\begin{equation}
+\rho_B = \text{Tr}_A(\rho_0) = \langle0|\rho_0|0\rangle_A + \langle1|\rho_0|1\rangle_A.
+\end{equation}
+
+The density matrices for these subsets can be used to compute the so-called von Neumann entropy, which is one of the possible measures of entanglement.
+A pure state has entropy equal to zero, while entangled states have an entropy larger than zero.
+The von Neumann entropy is defined as
+
+\begin{equation}
+S(A, B) = -\text{Tr}(\rho_{A,B} \log_2(\rho_{A,B})).
+\end{equation}
+
+We will use this definition, to measure the level of entanglement present in the system as a function of the interatction strength parameter $\lambda$.
 

### 2023-05-31 src/runs/vqe_1qubit.py
Generating figures
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -6,6 +6,7 @@ from src.ops import *
 from scipy.optimize import minimize
 import matplotlib.pyplot as plt
 from tqdm import tqdm
+import time
 
 # Solving the Hamiltonian analytically
 
@@ -22,6 +23,19 @@ def plot_energies(analytical, numerical=None, lmd=np.zeros(20)):
     plt.show()
 
 
+def plot_against_numerical(energies, numerical, lmbds):
+    figs, axs = plt.subplots(1, 1, figsize=(12, 12))
+    for i in range(2):
+        axs.plot(lmbds, numerical[:, i], label=f"$E_{i+1}$")
+    axs.scatter(lmbds, energies, label=f"$E_0$")
+    axs.set_xlabel(r"$\lambda$")
+    axs.set_label("Energy")
+    axs.grid()
+    axs.set_title("Energy levels as a function of interaction strength $\lambda$")
+    axs.legend()
+    plt.show()
+
+
 def plot_vqe_energies(energies, analytical, lmbds):
     figs, axs = plt.subplots(1, 1, figsize=(12, 12))
     for i in range(2):
@@ -30,10 +44,12 @@ def plot_vqe_energies(energies, analytical, lmbds):
     axs.set_xlabel(r"$\lambda$")
     axs.set_label("Energy")
     axs.grid()
+    axs.set_title("Energy levels as a function of interaction strength $\lambda$")
     axs.legend()
     plt.show()
 
 
+start_time = time.time()
 # lam = np.arange(0, 1, 0.05)
 lam = np.linspace(0.0, 1.0, 40)
 system_ener_an = np.zeros((len(lam), 2))
@@ -45,8 +61,10 @@ for i, l in enumerate(lam):
     perm = eig_vals.argsort()
     system_ener_an[i] = eig_vals[perm]
 
+print(f"Time taken for the analytical method: {time.time() - start_time}")
 # Solving the Hamiltonian numerically
 
+start_time = time.time()
 system_ener_num = np.zeros((len(lam), 2))
 for i, l in enumerate(lam):
     Ham, _ = hamiltonian_1qubit(
@@ -57,7 +75,10 @@ for i, l in enumerate(lam):
     perm = eig_vals.argsort()
     system_ener_num[i] = eig_vals[perm]
 
+print(f"Time taken for the numerical method: {time.time() - start_time}")
 
+
+start_time = time.time()
 num_shots = 1000
[…]

### 2023-05-31 doc/paper.tex
Figures for 2 qubit system
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 Prior to simulating the Lipkin model using our VQE alogorithm, we constructed two simple Hamiltonians each representing a system comprised of one and two qubits respectively.
-We started by setting up the one qubit system where the Hamiltonian matrix is given by the simple eqution constituting a non-interacting part $H_0$ and an interacting part $H_1$,
+We started with the one-body Hamiltonian, where we rewrote the Hamiltonian in terms of the Pauli operators and solved for the lowest eigenvalue using eigendecomposition and the power method before we set up our ansatz and use the VQE to find the ground state energy.
 

### 2023-05-31 src/runs/vqe_2qubit.py

@@ -6,31 +6,34 @@ from src.ops import *
 from scipy.optimize import minimize
 import matplotlib.pyplot as plt
 from tqdm import tqdm
+import time
 
 H = hamiltonian_2qubit(1.0, 2.0, 3.0, np.array([0.0, 2.5, 6.5, 7.0]))
 lmbds = np.linspace(0.0, 1.0, 50)
 eigvals_an = np.zeros((len(lmbds), H.shape[1]))
-eigvals_num1 = np.zeros((len(lmbds), H.shape[1]))
-eigvals_num2 = np.zeros((len(lmbds), H.shape[1]))
 entropy = np.zeros((len(lmbds), H.shape[1]))
 
+start_time = time.time()
 for i, lmbd in enumerate(lmbds):
     H = hamiltonian_2qubit(lmbd, 2.0, 3.0, np.array([0.0, 2.5, 6.5, 7.0]))
     eigvals, eigvecs = np.linalg.eig(H)
-    eig_num1, _ = power_iteration(H, 400)
-    eig_num2 = QR_solver(H)
     perm_an = eigvals.argsort()
-    perm_num1 = eig_num1.argsort()
-    perm_num2 = eig_num2.argsort()
     eigvals_an[i] = eigvals[perm_an]
-    eig_num1 = eig_num1[perm_num1]
-    eig_num2 = eig_num2[perm_num2]
-    eigvals_num1[i] = eig_num1
-    eigvals_num2[i] = eig_num2
     eigvecs = eigvecs[:, perm_an]
     entropy[i] = von_neumann(eigvals, eigvecs, 0)
+print(f"Time taken for the analytical method: {time.time() - start_time}")
 
+start_time = time.time()
+eigvals_num1 = np.zeros((len(lmbds), H.shape[1]))
+for i, lmbd in enumerate(lmbds):
+    H = hamiltonian_2qubit(lmbd, 2.0, 3.0, np.array([0.0, 2.5, 6.5, 7.0]))
+    eig_num1, _ = power_iteration(H, 400)
+    perm_num1 = eig_num1.argsort()
+    eig_num1 = eig_num1[perm_num1]
+    eigvals_num1[i] = eig_num1
+print(f"Time taken for the numerical method: {time.time() - start_time}")
 
+start_time = time.time()
 learning_rate = 0.2
 momentum = 0.2
 num_shots = 500
@@ -46,6 +49,7 @@ for i, lmbd in enumerate(tqdm(lmbds2)):
 # print(
[…]

### 2023-05-31 src/runs/vqe_2qubits.py

@@ -1,55 +1,0 @@
-from __future__ import annotations
-import numpy as np
-import math
-from scipy.optimize import minimize
-from tqdm import tqdm
-from src.solvers import *
-from src.ops import *
-
-basis0 = np.array([1.0, 0.0])
-basis00 = np.kron(basis0, basis0)
-basis10 = np.array([0.0, 1.0, 0.0, 0.0])
-basis01 = np.array([0.0, 0.0, 1.0, 0.0])
-basis11 = np.array([0.0, 0.0, 0.0, 1.0])
-
-# Testing if the initialization of a two qubit ansatz works correctly
-
-Z = np.array([[1.0, 0.0], [0.0, -1.0]])
-I = np.eye(2)
-II = np.kron(I, I)
-IZ = np.kron(I, Z)
-ZI = np.kron(Z, I)
-ZZ = np.kron(Z, Z)
-
-# Our goal is to rewrite the hamiltonian of the system in terms of the Pauli Z matrix, and
-# the identity matrix. Using this two projection operators below, we can write four equation
-# which we will simplify, which turn out to be equivalent to the standard projection operators.
-P_0 = (1 / 2) * (I + Z)
-P_1 = (1 / 2) * (I - Z)
-
-assert np.allclose(np.outer(basis00, basis00), (1 / 4) * (II + ZI + IZ + ZZ))
-assert np.allclose(np.outer(basis10, basis10), (1 / 4) * (II + ZI - IZ - ZZ))
-assert np.allclose(np.outer(basis01, basis01), (1 / 4) * (II - ZI + IZ - ZZ))
-assert np.allclose(np.outer(basis11, basis11), (1 / 4) * (II - ZI - IZ + ZZ))
-
-# However, instead of writing each of those by them selves, we are going to set
-# up a 4x4 matrix, and compute four coefficients A,B,C and D in order to rewrite
-# the Hamiltonian using tensor products consisting of I and Z.
-
-# print(measure_energy_2q())
-
-learning_rate = 0.2
-momentum = 0.2
-num_shots = 500
-max_epochs = 5000
-lmbds = np.linspace(0.0, 1.0, 11)
-min_energy = np.zeros(len(lmbds))
-epochs = np.zeros(len(lmbds))
-for i, lmbd in enumerate(tqdm(lmbds)):
-    angles = np.random.uniform(0.0, np.pi, size=4)
-    angles, epoch, energy, delta_energy = VQE_momentum(
[…]

### 2023-05-31 doc/paper.tex
Finished the report and code for project, need to add a couple of comments and references in the paper
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

--- 2023-05-31 doc/paper.tex
@@
 
-To assess the VQE's efficacy, we apply it to quantum systems representing single and two-particle systems, comparing its performance against numerical and analytical solvers.
+To assess the VQE's efficacy, we apply it to quantum systems representing single and two-particle systems, comparing its performance against an analytical solver.
 Transitioning to the Lipkin model, we leverage the properties of annihilation and creation operators to reformulate the Lipkin model's Hamiltonian first in terms of the quasi-spin operator and then Pauli operators.
@@
 
-\subsection{Power iteration}
-In our endeavor to showcase the prowess of the variational quantum algorithm, we will employ the power method for finding an approximate solution to the problem of finding the ground state of the Hamiltonian.
-However, we will tailor it slightly to seek the lowest eigenvalue instead of the largest.
-This method is remarkably straightforward, involving the computation of a matrix-vector product between our target matrix and a randomly initialized vector.
-The resulting product is then assigned to a new vector, which is subsequently normalized.
-This iterative process is repeated for a specified number of iterations or until convergence is attained.
-
-The key insight of the power method is that with each multiplication, the initially initialized vector aligns itself more closely with the eigenvector corresponding to the lowest eigenvalue of the target matrix.
-Convergence is determined by measuring the absolute value of the difference between the product of the current and previous iterations of the algorithm.
-
-By employing the power method with the necessary modifications, we can effectively find the lowest eigenvalue and its corresponding eigenvector, thereby gaining valuable insights into the properties and behavior of the matrix of interest.
-A detailed descripton of the method is presented in [Chapla, p.794].
-
 \subsection{Variational Quantum Eigensolver}
@@
 \subsection{Solving one and two qubit Hamiltonians} 
-Prior to simulating the Lipkin model using our VQE alogorithm, we constructed two simple Hamiltonians each representing a system comprised of one and two qubits respectively.
-We started by setting up the one qubit system where the Hamiltonian matrix is given by the simple eqution constituting a non-interacting part $H_0$ and an interacting part $H_1$,
+Prior to simulating the Lipkin model using our VQE algorithm, we constructed two simple Hamiltonians representing systems comprised of one and two qubits respectively.
+Initially, we focused on the one-body Hamiltonian and expressed it in terms of the Pauli operators.
+By performing eigendecomposition, we obtained the lowest eigenvalue as a reference.
+Subsequently, we set up our ansatz and employed the VQE algorithm to find the ground state energy.
+
+The figure below displays the resulting energies of the one-body Hamiltonian, with the VQE results indicated by dots in the lower plot.
+Notably, the VQE algorithm successfully approximates the ground state energy of this Hamiltonian.
+Examining the energy levels of the system, we observe that up until $\lambda=\frac{2}{3}$, the lowest eigenstate is predominantly characterized by $|0>$, while the upper eigenstate is dominated by $|1>$.
+However, beyond this threshold, a significant change occurs as the energy levels invert their direction, indicating an interchange in the character of the eigenvectors.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.9\textwidth, height=0.65\textwidth]{figures/VQE_vs_Analytical_1qubit.png}
+  \caption{Energy levels for the one-body Hamiltonian from subsection 3.2.1.}
+  \label{fig:ex2}
+\end{figure}
+
+When simulating the two-body Hamiltonian, we observe that the VQE again closely resembles the analytical solution, although the approximation has slighlty less precision than in the case of the one-body Hamiltonian above.
+The two lewest energy levels also display a similiar behavior to that of the energy levels above, albeit for different values of the interaction strength.
+The energy levels are displayed in teh figure below.
+Another thing we notice about the VQE, is the already apparent slow down of the computations, which is caused by our representation of the operators in terms of matrices.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.9\textwidth, height=0.75\textwidth]{figures/VQE_vs_Analytical_2qubits.png}
+  \caption{Energy levels for the two-body Hamiltonian from subsection 3.2.2.}
+  \label{fig:ex3}
+\end{figure}
+
+Besides the energy levels of each of the model's eigenstates, we have also measured the von Neumann entropy for the first qubit of the system in order to study entanglement within this particular system.
+As can be seen from Figure $\ref{fig:entropy}$, there is a jump in entropy around the point where we have a level crossing, at around $\lambda=0.4$.
+At an interaction strength of $\lambda=0$, the many-body states are solely defined by their computational basis states.
+As we increase the interaction strength, the degree of mixing increases, leading to a rise in entropy.
+This trend continues until we reach the level crossing point, where we observe a sudden and additional increase in entropy.
+We observe similar behaviors for the other states.
+The key takeaway from this example is that entanglement is influenced by the Hamiltonian itself, including the strength of the interaction matrix elements and the non-interacting energies.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.9\textwidth, height=0.75\textwidth]{figures/Von_Neumann_entropy_2qubit.png}
+  \caption{Von Neumann entropy for qubit 0 of the two-body Hamiltonian of subsection 3.2.2.}
+  \label{fig:entropy}
+\end{figure}
+
+\subsection{Results of simulating the Lipkin model}
+While simulating the Lipkin model, we observed that our VQE algorithm, written purely in numpy without the utilization of any optimization libraries, started to display worse levels of precision.
+As a result, we improved our VQE algorithm by utilizing the \textit{minimize} method of the \textit{Scipy} library, resulting in visibly better precision and faster convergence times.
+
+Starting with the case when the total spin of the Lipkin model is $J=1$, we see that compared to the previous Hamiltonians, the energies corresponding to the eigenstates of the system display symmetries about the x-axis.
+Specifically, the lowest and highest eigenvalues are antisymmetric but proportional.
+In other words, as $E_2$ becomes larger as a function of $V/\epsilon$, $E_0$ increases in the opposite (negative) direction.
+Interestingly, the middle eigenvalue does not evolve and remains equal to zero at all times, correctly indicating the state where each of the two fermions occupies lower and upper energy levels, respectively.
+The results recorded during simulation of the the Lipkin model with two fermions are shown in Figure $\ref{fig:lipkin_J1}$.
+
+\begin{figure}[H]
+  \centering
+	\includegraphics[width=0.9\textwidth, height=0.75\textwidth]{figures/VQE_vs_Analytical_Lipkin_J1_Scipy.png}
+	\caption{Energy levels of the Lipkin model with a total spin J=1. Solution obtained using the VQE algorithm are marked with 'x'.}
+  \label{fig:lipkin_J1}
+\end{figure}
+
+We observe similar symmetries when the total spin of the Lipkin model is $J=2$.
+In this case, the eigenvalues above the x-axis perfectly mirror the eigenvalues below.
+These symmetries arise from the conservation of total spin or angular momentum, as the Lipkin model's Hamiltonian remains invariant under spin rotations.
+This invariance gives rise to symmetries in the energy levels and eigenstates of the system.
+The preservation of total spin leads to specific patterns and symmetries, such as the eigenvalues being antisymmetric but proportional, and the middle eigenvalue remaining unchanged.
+These symmetries are rooted in the fundamental preservation of total spin, which governs the behavior of the Lipkin model.
+
+Moving on to the approximation of the lowest eigenvalue, we were once again able to achieve a good fit to the analytical solution (see Fig.
+$\ref{fig:lipkin_J2}$, albeit at the expense of time.
+The algorithm required approximately 200-300 seconds to complete the computations.
+This highlights the point that while classical computers can be used to simulate algorithms like VQE and quantum systems with a few qubits, the increase in degrees of freedom associated with larger simulations makes it increasingly challenging.
+Simulating systems with more than four qubits becomes difficult, and simulating systems with more than ten qubits becomes nearly impossible, especially on standard laptops.
+Therefore, the limitations of classical computers in simulating large quantum systems underscore the need for quantum computers and algorithms like VQE to fully explore and understand complex quantum phenomena.
+
+\begin{figure}[H]
+  \centering
+	\includegraphics[width=0.9\textwidth, height=0.75\textwidth]{figures/VQE_vs_Analytical_Lipkin_J2_Scipy.png}
+	\caption{Energy levels of the Lipkin model with a total spin J=2. Solution obtained using the VQE algorithm are marked with 'o'.}
+  \label{fig:lipkin_J2}
+\end{figure}
+
+\subsection{Summary}
+We initiated this project by examining the behavior of quantum gates on individual quantum bits and investigating their combination to construct more intricate circuits applicable to an arbitrary number of particles.
+Subsequently, we proceeded to implement the Variational Quantum Eigensolver (VQE) and develop the necessary circuits to prepare suitable ansatzes, which we coupled with the VQE to approximate the ground state energy of three distinct Hamiltonians.
+Beginning with a single-body Hamiltonian and culminating with the Lipkin model consisting of four fermions, we evaluated the performance of our VQE algorithm against an analytical solver, thereby gaining valuable insights.
+
+Based on our attained results, we can assert that VQE approximates the lowest eigenvalue of the presented Hamiltonian matrices with good precision.
+However, it is susceptible to the curse of dimensionality stemming from the nature of qubits and the systems they compose.
+While the algorithm operates swiftly for systems encompassing one or two qubits, a significant reduction in computational speed becomes evident even for systems involving four particles.
+Consequently, further efforts must be directed toward developing more potent quantum computers.
+As of now, conventional analytical solvers such as eigendecomposition still outperform VQE on classical computers, and as such, remain a preferable choice of algorithm.
+
+\subsection{References}
 

### 2023-05-31 src/runs/lipkinGM_model_J1.py

@@ -3,48 +3,36 @@ from src.solvers import *
 from src.ops import *
 from tqdm import tqdm
 import matplotlib.pyplot as plt
+import time
+from scipy.optimize import minimize
 
 # The case of spin J=1, solving for eigenvalues using analytical and numerical methods
 
 
-def plot_energy(energy_an, energy_vqe, v_values_ana, v_values_vqe):
+def plot_energy(energy_an, energy_vqe, energy_vqe_scipy, v_values_ana, v_values_vqe):
     fig, axs = plt.subplots(1, 1, figsize=(12, 12))
     for i in range(len(energy_an[0])):
         axs.plot(v_values_ana, energy_an[:, i], label=f"$E_{i}$")
     axs.set_xlabel(r"$V/\epsilon$")
-    axs.scatter(v_values_vqe, energy_vqe, label="VQE Energy Scipy", marker="o")
+    axs.scatter(v_values_vqe, energy_vqe, label="VQE Energy", marker="o")
+    axs.scatter(v_values_vqe, energy_vqe_scipy, label="VQE Energy Scipy", marker="x")
     axs.set_ylabel(r"$E/\epsilon$")
+    axs.set_title("Energy levels as a function of interaction parameter V$")
     axs.legend(loc="upper left")
     plt.show()
 
 
-def plot_energy_FH(energy_an, energy_an_FH, v_values_ana):
-    fig, axs = plt.subplots(1, 1, figsize=(12, 12))
-    for i in range(len(energy_an[0])):
-        # axs.plot(v_values_ana, energy_an[:, i], label=f"$E_{i}$")
-        axs.plot(v_values_ana, energy_an_FH[:, i], label=f"$E_FH{i}$")
-    axs.set_xlabel(r"$V/\epsilon$")
-    axs.set_ylabel(r"$E/\epsilon$")
-    axs.legend(loc="upper left")
-    plt.show()
-
-
+start_time = time.time()
 Vs_an = np.linspace(0.0, 2.0, 200)
 eigvals_an = np.zeros((len(Vs_an), 3))
-eigvals_num_PI = np.zeros((len(Vs_an), 3))
-eigvals_num_QR = np.zeros((len(Vs_an), 3))
 
 for i, v in enumerate(tqdm(Vs_an)):
     H = lipkin_H_J1(v)
     eigvals, eigvecs = np.linalg.eig(H)
-    eig_num_PI, _ = power_iteration(H, 1000)
-    eig_num_QR = QR_solver(H, iters=1000)
[…]

### 2023-05-31 src/runs/lipkinMG_model_J2.py

@@ -3,31 +3,18 @@ from src.ops import *
 import numpy as np
 from tqdm import tqdm
 import matplotlib.pyplot as plt
-
+import time
 
 v_values_an = np.linspace(0, 2.0, 100)
-w_values_an = np.linspace(0, 1.0, 100)
 eigvals_an = np.zeros((len(v_values_an), 16))
-entropy = np.zeros((len(v_values_an), 16))
 
+start_time = time.time()
 for i, v in enumerate(tqdm(v_values_an)):
     H = lipkin_H_J2_Pauli(v)
-    HF = lipkin_H_J2_Pauli(v, w_values_an[i], True)
-
     eig_vals, eig_vecs = np.linalg.eig(H)
     eig_perm = eig_vals.argsort()
     eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]
-
-print(eigvals_an[49][0])
-
-
-# fig, axs = plt.subplots(1, 1, figsize=(8, 8))
-# for i in range(len(eigvals_an[0])):
-#     axs.plot(v_values_an, eigvals_an[:, i], label=f"$E_{i}$", linestyle="dashed")
-# axs.set_xlabel(r"$V/\epsilon$")
-# axs.set_ylabel(r"$E/\epsilon$")
-# axs.legend(loc="upper left")
-# plt.show()
+print(f"Time taken for the analytical method: {time.time() - start_time}")
 
 # In order to be able to perform the measurement in the Z basis, we will need to
 # rotate our measurement basis into the ZIII basis. In order to do so, we are going
@@ -38,24 +25,20 @@ print(eigvals_an[49][0])
 # J=2, indicating four fermions.
 
 
-def get_gradient(angles, v, number_shots, unitaries):
-    grad = np.zeros(len(angles))
-    for index, angle in enumerate(angles):
-        tmp = angles.copy()
-        tmp[index] += np.pi / 2
-        energy_plus = measure_energy_mul(tmp, v, number_shots)
-        tmp[index] -= np.pi
-        energy_minus = measure_energy_mul(tmp, v, number_shots)
-        grad[index] = (energy_plus - energy_minus) / 2
-    return grad
-
-
+start_time = time.time()
 unitaries = prep_circuit_lipkin_J2()
-
 v_vals = np.linspace(0.0, 2.0, 8)
-
 ener = VQE_scipy(
[…]

### 2023-05-31 doc/paper.tex
Finished report
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 The value of the parameter $V_{12}$ represents the strength of the coupling between the two states.
-([Fys5419, project 1])
+([Morten, project 1])
 
@@
 
-\subsection{References}
+\section{References}
+
+[Hundt, 2022] Hundt Robert (2022). Qunatum Computing for Programmers. Cambridge University Press.
+
+[Nielsen, 2021] Nielsen Michael (2021). Qunatum Computation adm Quantum Informantion. Cambridge University Press.
+
+[Physical review C, 2022] Manqoba Q. Hlatshwayo (2022). Simulating excited states og the Lipkin model on a Quantum Computer.
+
+[Basdevant, 2023] Basdevant Jean-Louis (2023). Lectures on Quantum Mechanics With Problems, Exercises and Solutions. Springer Nature Switzerland.
+
+[Scherer, 2019] Scherer Wolfgang (2019). Mathematics of Quantum Computing an Introduction. Springer Nature Switzerland.
+
+[Morten Hjorth-Jensen, 2023] M. Hjorth-Jensen, ”Week 11: VQE and project discussion,” URL: https://github.com/CompPhysics/QuantumComputingMachineLearning/tree/gh-pages/doc/pub/week11 April 2023.
+
+[Morten Hjorth-Jensen, 2023] M.
+Hjorth-Jensen, ”Week 8: Solving quantum mechanical problems,” URL: https://github.com/CompPhysics/QuantumComputingMachineLearning/tree/gh-pages/doc/pub/week8, March 2023.
+
+[Morten Hjorth-Jensen, 2023] M.
+Hjorth-Jensen, ”Project 1: Quantum Computing and many-body problems,” URL: https://github.com/CompPhysics/QuantumComputingMachineLearning/blob/gh-pages/doc/Projects/2023/Project1/pdf/Project1.pdf March 2023.
 

### 2023-06-08 src/runs/lipkinMG_model_J2.py
Start work on fixing typos, adding full hamiltonian section, and polishing the code
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -10,7 +10,7 @@ eigvals_an = np.zeros((len(v_values_an), 16))
 
 start_time = time.time()
 for i, v in enumerate(tqdm(v_values_an)):
-    H = lipkin_H_J2_Pauli(v)
+    H = lipkin_H_J2_Pauli(v, 0, full=True)
     eig_vals, eig_vecs = np.linalg.eig(H)
     eig_perm = eig_vals.argsort()
     eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]
@@ -39,6 +39,6 @@ for i in range(len(eigvals_an[0])):
 # axs.scatter(v_vals, ener, label="VQE Energy Scipy", marker="o")
 axs.set_xlabel(r"$V/\epsilon$")
 axs.set_ylabel(r"$E/\epsilon$")
-axs.set_title("Energy levels as a function of interaction parameter V$")
+axs.set_title("Energy levels as a function of interaction parameter W")
 axs.legend(loc="upper left")
 plt.show()

### 2023-06-08 src/solvers.py

@@ -485,7 +485,7 @@ def measure_energy_2q(
     return coeffs[0] + np.sum(consts * exp_vals) / shots
 
 
-def measure_energy_mul(angles, v, num_shots, w=0.0):
+def measure_energy_mul(angles, v, num_shots):
     gates = prep_circuit_lipkin_J2()
     state4 = ansatz_4qubit(angles)
     measurements = np.zeros((len(gates), num_shots))
@@ -498,6 +498,7 @@ def measure_energy_mul(angles, v, num_shots, w=0.0):
     consts = np.concatenate(
         (0.5 * np.ones(4), -0.5 * v * np.ones(6), 0.5 * v * np.ones(6))
     )
+     
     for i in range(len(exps)):
         counts = [len(np.where(measurements[i] == j)[0]) for j in range(16)]
         for out, count in enumerate(counts):

### 2023-06-09 src/runs/lipkinMG_model_J2.py
Code enabling computation of the full Lipkin added, although the number operator may still be wrong
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -10,7 +10,7 @@ eigvals_an = np.zeros((len(v_values_an), 16))
 
 start_time = time.time()
 for i, v in enumerate(tqdm(v_values_an)):
-    H = lipkin_H_J2_Pauli(v, 0, full=True)
+    H = lipkin_H_J2_Pauli(v)
     eig_vals, eig_vecs = np.linalg.eig(H)
     eig_perm = eig_vals.argsort()
     eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]
@@ -28,15 +28,24 @@ print(f"Time taken for the analytical method: {time.time() - start_time}")
 start_time = time.time()
 unitaries = prep_circuit_lipkin_J2()
 v_vals = np.linspace(0.0, 2.0, 8)
+# w_vals = np.linspace(0.0, 2.0, 8)
 ener = VQE_scipy(
-    measure_energy_mul, v_vals, 5 * 8, 1000, unitaries, 0.0, 2 * np.pi, "BFGS"
+    measure_energy_mul,
+    v_vals,
+    5 * 8,
+    1000,
+    0.0,
+    2 * np.pi,
+    "BFGS",
+    # w_vals,
+    # full=True,
 )
 print(f"Time taken for the VQE method: {time.time() - start_time}")
 
 fig, axs = plt.subplots(1, 1, figsize=(8, 8))
 for i in range(len(eigvals_an[0])):
     axs.plot(v_values_an, eigvals_an[:, i], label=f"$E_{i}$")
-# axs.scatter(v_vals, ener, label="VQE Energy Scipy", marker="o")
+axs.scatter(v_vals, ener, label="VQE Energy Scipy", marker="o")
 axs.set_xlabel(r"$V/\epsilon$")
 axs.set_ylabel(r"$E/\epsilon$")
 axs.set_title("Energy levels as a function of interaction parameter W")

### 2023-06-09 src/solvers.py

@@ -94,7 +94,7 @@ def lipkin_H_J2_Pauli(v, w=0, full=False):
     IYIY = np.kron(I, np.kron(Y, np.kron(I, Y)))
     IIYY = np.kron(I, np.kron(I, np.kron(Y, Y)))
 
-    N1 = np.kron((1 / 2)*(I - Z), np.kron(I, np.kron(I, I)))
+    N1 = np.kron((1 / 2) * (I - Z), np.kron(I, np.kron(I, I)))
     N2 = np.kron(I, np.kron((1 / 2) * (I - Z), np.kron(I, I)))
     N3 = np.kron(I, np.kron(I, np.kron((1 / 2) * (I - Z), I)))
     N4 = np.kron(I, np.kron(I, np.kron(I, (1 / 2) * (I - Z))))
@@ -210,7 +210,6 @@ def QR_solver(H, delta=1e-6, iters=100):
 
     for iteration in range(iters):
         Q, R = np.linalg.qr(H)
-        # A = np.dot(R, Q)
         H = Q.T @ H @ Q
 
         # Check convergence
@@ -485,7 +484,7 @@ def measure_energy_2q(
     return coeffs[0] + np.sum(consts * exp_vals) / shots
 
 
-def measure_energy_mul(angles, v, num_shots):
+def measure_energy_mul(angles, v, num_shots, w=None, full=False):
     gates = prep_circuit_lipkin_J2()
     state4 = ansatz_4qubit(angles)
     measurements = np.zeros((len(gates), num_shots))
@@ -495,10 +494,21 @@ def measure_energy_mul(angles, v, num_shots):
         measurements[i] = measurement
 
     exps = np.zeros(len(measurements))
-    consts = np.concatenate(
-        (0.5 * np.ones(4), -0.5 * v * np.ones(6), 0.5 * v * np.ones(6))
+
+    if full:
+        consts = np.concatenate(
+            (
+                0.5 * np.ones(4),
+                -0.5 * v * np.ones(6),
+                0.5 * v * np.ones(6),
+                0.5 * w * np.ones(12),
+            )
+        )
+    else:
+        consts = np.concatenate(
+            (0.5 * np.ones(4), -0.5 * v * np.ones(6), 0.5 * v * np.ones(6))
     )
-     
+    
     for i in range(len(exps)):
         counts = [len(np.where(measurements[i] == j)[0]) for j in range(16)]
         for out, count in enumerate(counts):
[…]

### 2023-06-09 tests/lipkin_W_part_test.py

@@ -6,21 +6,34 @@ from tqdm import tqdm
 
 np.set_printoptions(linewidth=200)
 
+zero = np.zeros((2, 2))
+I = Identity()
+Z = PauliZ()
+init_angles = np.random.uniform(0.0, 2 * np.pi, size=5 * 8)
+state4 = ansatz_4qubit(init_angles)
+
 H = lipkin_H_J2_Pauli(v=1.0, w=1.0, full=True)
 v_values_an = np.linspace(0, 2.0, 100)
 w_values_an = np.linspace(0, 1.0, 100)
 eigvals_an = np.zeros((len(v_values_an), 16))
 
+N1 = np.kron((1 / 2) * (I - Z), np.kron(I, np.kron(I, I)))
+N2 = np.kron(I, np.kron((1 / 2) * (I - Z), np.kron(I, I)))
+N3 = np.kron(I, np.kron(I, np.kron((1 / 2) * (I - Z), I)))
+N4 = np.kron(I, np.kron(I, np.kron(I, (1 / 2) * (I - Z))))
+N = N1 + N2 + N3 + N4
 
-for i, v in enumerate(tqdm(v_values_an)):
-    H = lipkin_H_J2_Pauli(v, w_values_an[i], True)
-
-    eig_vals, eig_vecs = np.linalg.eig(H)
-    eig_perm = eig_vals.argsort()
-    eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]
-
-circuits = prep_circuit_lipkin_J2()
-print(circuits)
+Nstate4 = N @ state4
+print(N)
+# for i, v in enumerate(tqdm(v_values_an)):
+#     H = lipkin_H_J2_Pauli(v, w_values_an[i], True)
+#
+#     eig_vals, eig_vecs = np.linalg.eig(H)
+#     eig_perm = eig_vals.argsort()
+#     eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]
+#
+# circuits = prep_circuit_lipkin_J2()
+# print(circuits)
 # fig, axs = plt.subplots(1, 1, figsize=(8, 8))
 # for i in range(len(eigvals_an[0])):
 #     axs.plot(v_values_an, eigvals_an[:, i], label=f"$E_{i}$")

### 2023-06-09 doc/paper.tex
Fixing som typos
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@
 
-\title{Simulating the Lipkin-Meshov-Glikov model on a classical computer}
+\title{Simulating the Lipkin-Meshkov-Glikov model on a classical computer}
 
@@
 Initially, we explore the outcomes arising from the application of diverse single qubit quantum gates to one qubit systems prepared in basis states.
-Subsequently, an entangler circuit is constructed, and one of the four bell pairs is prepared, to which we apply the Hadamard and CNOT gates.
+Following that, an entangler circuit is constructed, and one of the four bell pairs is prepared, to which we apply the Hadamard and CNOT gates.
 Utilizing our newfound grasp of the foundational principles governing quantum systems, we construct quantum circuits for generating ansatz states and methodically develop the VQE from ground up.
 
-To assess the VQE's efficacy, we apply it to quantum systems representing single and two-particle systems, comparing its performance against an analytical solver.
+To assess the VQE's efficiency, we apply it to quantum systems representing single and two-particle systems, comparing its performance against an analytical solver.
 Transitioning to the Lipkin model, we leverage the properties of annihilation and creation operators to reformulate the Lipkin model's Hamiltonian first in terms of the quasi-spin operator and then Pauli operators.
-Employing this encoding scheme, we apply the VQE to systems with spin J=1 and 2, corresponding to Lipkin systems with two and four fermions, respectively.
-We compare the results against exact solutions obtained through standard diagonalization methods.
+Employing this encoding scheme, we apply the VQE to the Lipkin model with spin J=1 and 2, corresponding to situation when the model comprises two and four fermions, respectively.
+We compare the results against exact solutions obtained through standard diagonalization method.
 This analysis provides valuable insights into the performance and accuracy of the VQE approach in tackling the Lipkin model's Hamiltonian.
@@
 However, it is worth noting that quantum computers still face challenges, such as larger error rates compared to classical computers and vulnerability to noise, which can make working with them potentially difficult.
-Although the field of quantum computing is expanding rapidly, widespread accessibility remains limited, with companies like IBM offering only highly restricted access to their quantum technology.
+Although the field of quantum computing is expanding rapidly, widespread accessibility remains limited, with companies like IBM offering only restricted access to their quantum technology.
 In light of these factors, we will employ a classical computer to simulate and analyze the simplified Lipkin-Meshkov-Glick model, showcasing the potential that quantum computing holds while acknowledging the current limitations in accessibility.
@@
 The primary distinction between classical computers and their quantum counterparts lies in the nature of the data they process.
-In classical computing, the fundamental data unit is the binary digit, commonly known as a bit, which represents different voltage levels associated with low and high states.
-On the other hand, quantum computers utilize particles, such as atoms, as the information unitm leading to a fundamentally different approach to data handling.
+On a classical computer, the fundamental data unit is the binary digit, commonly known as a bit, which represents different voltage levels associated with low and high states, or 0 and 1 as is also common.Quantum computers on the other hand, utilize particles as the information unit, hence leading to a fundamentally different approach to data handling.
 The unique properties of atoms and the underlying principles of quantum mechanics prevent us from treating them in the same manner as classical bits.
@@
 This implies that the operations applied to quantum systems are inherently reversible.
-Consequently, at any given point in time, a quantum system can be fully characterized by a set of linear unitary operators.
 As a result, data processing on a quantum computer necessitates the use of reversible operators, enabling the "uncomputation" of the outcomes produced by quantum gates.
@@
 		The action of the Pauli Y gate can be described as the mapping of $|0>$ to $i|1>$ and $|1>$ to $i¦0>$.
-The action of the gate can also be viewed as a rotation of the state of the qubit around the y-axis of the Bloch sphere by 180 degrees.
+The action of the gate can also be viewed as a rotation of the state of the qubit around the y-axis of the Bloch sphere (\textit{see Fig.1}) by 180 degrees.
 Presented below is the matrix representation of the quantum gate the result of applying it to a single qubit:
@@
 	\item[\textbf{IV.}] \textbf{Pauli-Z gate} \\
-		The Pauli Z gate flips the phase of $|1>$ to $-|1>$ while leaving $|0>$, meaning it flips the relative phase of the qubit.
+		The Pauli Z gate flips the phase of $|1>$ to $-|1>$ while leaving $|0>$ unchanged, meaning it flips the relative phase of the qubit.
 Presented below is the matrix representation of the quantum gate the result of applying it to a single qubit:
@@
 The name of the gate provides a clue about its functionality.
-By using the first qubit as a control signal, the CNOT gate flips the amplitudes of the second qubit if the control qubit is in the basis state |1>, which corresponds to an active high signal in classical computing.
-However, if the control qubit is in the state |0>, no changes are applied to the second qubit.
+By using the first qubit as a control signal, the CNOT gate flips the amplitudes of the second qubit if the control qubit is in the basis state $|1>$, which corresponds to an active high signal in classical computing.
+However, if the control qubit is in the state $|0>$, no changes are applied to the second qubit.
 The specific gate we are discussing here is the $CNOT_{0,1}$-gate, where the first qubit serves as the control signal.
@@
 This allows us to rewrite the Hamiltonian of the Lipkin model in a more convenient form.
-Consequently, we can express the Hamiltonian matrix in terms of the Pauli operators mentioned in the previous subsection.
+Consequently, we can express the Hamiltonian matrix in terms of the Pauli operators mentioned in the previous subsection, just as we did with the one- and two-qubit Hamiltonian.
 This transformation facilitates the analysis and study of the Lipkin model, providing insights into its energy levels and dynamics.

### 2023-06-09 src/runs/lipkinMG_model_J2.py
Finishing project
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -10,7 +10,7 @@ eigvals_an = np.zeros((len(v_values_an), 16))
 
 start_time = time.time()
 for i, v in enumerate(tqdm(v_values_an)):
-    H = lipkin_H_J2_Pauli(v)
+    H = lipkin_H_J2_Pauli(v, v, full=True)
     eig_vals, eig_vecs = np.linalg.eig(H)
     eig_perm = eig_vals.argsort()
     eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]
@@ -28,7 +28,7 @@ print(f"Time taken for the analytical method: {time.time() - start_time}")
 start_time = time.time()
 unitaries = prep_circuit_lipkin_J2()
 v_vals = np.linspace(0.0, 2.0, 8)
-# w_vals = np.linspace(0.0, 2.0, 8)
+w_vals = np.linspace(0.0, 2.0, 8)
 ener = VQE_scipy(
     measure_energy_mul,
     v_vals,

### 2023-06-09 src/solvers.py

@@ -7,6 +7,17 @@ from scipy.optimize import minimize
 from src.ops import *
 from tqdm import tqdm
 
+
+"""
+This file contains all the algorithms used for approximating/solving the 
+eigenvalue problems associated with each of the scripts found in the "runs"
+directory, starting with the extremely simple one-body Hamiltonian matrix, 
+and ending at the Lipkin model with spin J=2. It also contains the methods
+for the computation of the von neumann entropy for the two-body hamiltonian 
+(not Lipkin model with J=1). The code here contains little comments, however 
+the names should indicate which problem the methods correspond to.
+"""
+
 _PAULI_X = PauliX()
 _PAULI_Y = PauliY()
 _PAULI_Z = PauliZ()
@@ -485,6 +496,8 @@ def measure_energy_2q(
 
 
 def measure_energy_mul(angles, v, num_shots, w=None, full=False):
+
+    """Method used for measurements of expectation values """
     gates = prep_circuit_lipkin_J2()
     state4 = ansatz_4qubit(angles)
     measurements = np.zeros((len(gates), num_shots))
@@ -524,7 +537,11 @@ def measure_energy_mul(angles, v, num_shots, w=None, full=False):
         N3 = np.kron(I, np.kron(I, np.kron((1 / 2) * (I - Z), I)))
         N4 = np.kron(I, np.kron(I, np.kron(I, (1 / 2) * (I - Z))))
         N = N1 + N2 + N3 + N4
-        return np.sum(np.hstack((consts[:4] * exps[:4], consts[4:16] * exps[4:] + consts[16:] * exps[4:]))) / num_shots - 0.5*w*N[15,15]
+
+        # The following computation is used for the approximation of the ground state of a Hamiltonian 
+        # with W > 0. We should have computed the expectation value of the number operator, however we know 
+        # that in this case it should be 4, thus we directly include it in the computations 
+        return np.sum(np.hstack((consts[:4] * exps[:4], consts[4:16] * exps[4:] + consts[16:] * exps[4:]))) / num_shots -w*0.5*N[15,15]
     else:
         return np.sum(consts * exps) / num_shots
 
@@ -545,6 +562,11 @@ def chose_measurement(length, J=0):
 
 # Generalized version of VQE
 def VQE(eta, epochs, num_shots, init_angles, lmbd, J=0):
+    """ 
+    One of the two main version of the VQE alogrithm used for approximating the 
+    ground state energy of the hamiltonians presented in this project.
[…]

### 2023-06-09 tests/lipkin_W_part_test.py

@@ -9,8 +9,6 @@ np.set_printoptions(linewidth=200)
 zero = np.zeros((2, 2))
 I = Identity()
 Z = PauliZ()
-init_angles = np.random.uniform(0.0, 2 * np.pi, size=5 * 8)
-state4 = ansatz_4qubit(init_angles)
 
 H = lipkin_H_J2_Pauli(v=1.0, w=1.0, full=True)
 v_values_an = np.linspace(0, 2.0, 100)
@@ -23,8 +21,20 @@ N3 = np.kron(I, np.kron(I, np.kron((1 / 2) * (I - Z), I)))
 N4 = np.kron(I, np.kron(I, np.kron(I, (1 / 2) * (I - Z))))
 N = N1 + N2 + N3 + N4
 
-Nstate4 = N @ state4
 print(N)
+
+num_shots = 1000
+expectation = 0.0
+state4 = np.random.uniform(0, 1, 16)
+for i in range(num_shots):
+    init_angles = np.random.uniform(0.0, 2 * np.pi, size=1 * 8)
+    state4 = ansatz_4qubit(init_angles)
+    Nstate4 = N @ state4
+    expectation += np.sum(np.abs(Nstate4) ** 2)
+
+print(expectation / num_shots)
+
+
 # for i, v in enumerate(tqdm(v_values_an)):
 #     H = lipkin_H_J2_Pauli(v, w_values_an[i], True)
 #

### 2023-06-09 src/runs/bell_task_a.py
Project finished
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -5,6 +5,11 @@ from src.solvers import *
 from src.ops import *
 from src.bell_states import *
 
+"""
+This file presents the computations which are performed on bell states as is 
+asked in part A of the project
+"""
+
 if __name__ == "__main__":
     # Preparing bell states
     b00 = bell_state(0, 0)

### 2023-06-09 src/runs/gates_task_a.py

@@ -4,6 +4,11 @@ import math
 from src.solvers import *
 from src.ops import *
 
+"""
+This script presents the action of single qubit gates on qubits prepared in either state 
+|0> or state |1>. 
+"""
+
 if __name__ == "__main__":
     # Preparing two seperate qubits each in one of the basis state |0> and |1>
     psi0 = state(alpha=1.0)

### 2023-06-09 src/runs/lipkinGM_model_J1.py

@@ -6,7 +6,7 @@ import matplotlib.pyplot as plt
 import time
 from scipy.optimize import minimize
 
-# The case of spin J=1, solving for eigenvalues using analytical and numerical methods
+# The case of spin J=1, solving for eigenvalues using analytical and numerical methos
 
 
 def plot_energy(energy_an, energy_vqe, energy_vqe_scipy, v_values_ana, v_values_vqe):

### 2023-06-09 src/runs/lipkinMG_model_J2.py

@@ -5,6 +5,9 @@ from tqdm import tqdm
 import matplotlib.pyplot as plt
 import time
 
+# solving for the ground state energy of the lipkin Hamiltonian of the system with total spin J=2,
+# meaning that the system conatains four fermions which are spread across two energy levels.
+
 v_values_an = np.linspace(0, 2.0, 100)
 eigvals_an = np.zeros((len(v_values_an), 16))
 

### 2023-06-09 src/runs/vqe_1qubit.py

@@ -8,6 +8,11 @@ import matplotlib.pyplot as plt
 from tqdm import tqdm
 import time
 
+"""
+THis file contains both the analytical solution of the eigenvalue problem associated with the 
+one-body hamiltonian presented in part B and C of the project, as well as the VQE algorithm 
+using our own gradient computations for the minimization of the expectation value"""
+
 # Solving the Hamiltonian analytically
 
 

### 2023-06-09 src/runs/vqe_2qubit.py

@@ -8,6 +8,12 @@ import matplotlib.pyplot as plt
 from tqdm import tqdm
 import time
 
+
+"""
+THis file contains both the analytical solution of the eigenvalue problem associated with the 
+two-body hamiltonian presented in part D and E of the project, as well as the VQE algorithm 
+using our own gradient computations for the minimization of the expectation value"""
+
 H = hamiltonian_2qubit(1.0, 2.0, 3.0, np.array([0.0, 2.5, 6.5, 7.0]))
 lmbds = np.linspace(0.0, 1.0, 50)
 eigvals_an = np.zeros((len(lmbds), H.shape[1]))

### 2023-06-09 src/solvers.py

@@ -497,7 +497,9 @@ def measure_energy_2q(
 
 def measure_energy_mul(angles, v, num_shots, w=None, full=False):
 
-    """Method used for measurements of expectation values """
+    """Method used for measurements of expectation values used for computations associated 
+    with the lipkin model only
+    """
     gates = prep_circuit_lipkin_J2()
     state4 = ansatz_4qubit(angles)
     measurements = np.zeros((len(gates), num_shots))
@@ -530,18 +532,9 @@ def measure_energy_mul(angles, v, num_shots, w=None, full=False):
             elif out > 7:
                 exps[i] -= count
     if full:
-        I = np.eye(2)
-        Z = PauliZ()
-        N1 = np.kron((1 / 2) * (I - Z), np.kron(I, np.kron(I, I)))
-        N2 = np.kron(I, np.kron((1 / 2) * (I - Z), np.kron(I, I)))
-        N3 = np.kron(I, np.kron(I, np.kron((1 / 2) * (I - Z), I)))
-        N4 = np.kron(I, np.kron(I, np.kron(I, (1 / 2) * (I - Z))))
-        N = N1 + N2 + N3 + N4
-
         # The following computation is used for the approximation of the ground state of a Hamiltonian 
-        # with W > 0. We should have computed the expectation value of the number operator, however we know 
-        # that in this case it should be 4, thus we directly include it in the computations 
-        return np.sum(np.hstack((consts[:4] * exps[:4], consts[4:16] * exps[4:] + consts[16:] * exps[4:]))) / num_shots -w*0.5*N[15,15]
+        # with W > 0.
+        return np.sum(np.hstack((consts[:4] * exps[:4], consts[4:16] * exps[4:] + consts[16:] * exps[4:]))) / num_shots -w*0.5*4*np.abs(state4)**2 # expectation value of the number operator 
     else:
         return np.sum(consts * exps) / num_shots
 

### 2023-06-09 src/runs/lipkinMG_model_J2.py
Finished
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -13,7 +13,7 @@ eigvals_an = np.zeros((len(v_values_an), 16))
 
 start_time = time.time()
 for i, v in enumerate(tqdm(v_values_an)):
-    H = lipkin_H_J2_Pauli(v, v, full=True)
+    H = lipkin_H_J2_Pauli(v)
     eig_vals, eig_vecs = np.linalg.eig(H)
     eig_perm = eig_vals.argsort()
     eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]

### 2023-06-09 tests/init_2.py

@@ -1,0 +1,20 @@
+import numpy as np
+from src.solvers import *
+from src.ops import *
+import matplotlib.pyplot as plt
+
+# np.random.seed(271)
+# np.random.seed(1338)
+np.random.seed(9)
+np.set_printoptions(linewidth=200)
+angles = np.random.uniform(low=0.0, high=2 * np.pi, size=(1, 8))
+
+
+eta = 0.002
+mnt = 0.02
+epochs = 1000
+num_shots = 1000
+
+_, _, energy, _ = VQE(eta, epochs, num_shots, angles, 1.0, J=2)
+
+print(energy)

### 2023-06-09 tests/init_4qubit.py

@@ -1,0 +1,318 @@
+import numpy as np
+from src.solvers import *
+from src.ops import *
+import matplotlib.pyplot as plt
+
+# np.random.seed(271)
+# np.random.seed(1338)
+np.random.seed(9)
+np.set_printoptions(linewidth=200)
+angles = np.random.uniform(low=0.0, high=2 * np.pi, size=(1, 8))
+
+
+# Check if the init of a 4 qubit state works fine
+state4 = ansatz_4qubit(angles)
+
+# gates = prep_circuit_lipkin_J2()
+
+# vals = measure_energy_mul(angles, num_shots=1000, v=1.0)
+SWAP, CNOT10, H, S, I = Swap(), Cnot(1, 0), Hadamard(), Sgate(), Identity()
+Sdag = S.conj().T
+
+
+measures = []
+unitaries = []
+U_ZIII = np.kron(I, np.kron(I, np.kron(I, I)))
+unitaries.append(U_ZIII)
+U_IZII = np.kron(np.kron(I, I) @ SWAP, np.kron(I, I))
+unitaries.append(U_IZII)
+U_IIZI = U_IZII @ np.kron(I, np.kron(np.kron(I, I) @ SWAP, I))
+unitaries.append(U_IIZI)
+U_IIIZ = U_IIZI @ np.kron(I, np.kron(I, np.kron(I, I) @ SWAP))
+unitaries.append(U_IIIZ)
+
+U_ZIZI = np.kron(CNOT10, np.kron(I, I)) @ np.kron(I, np.kron(SWAP @ np.kron(I, I), I))
+
+U_XX = CNOT10 @ np.kron(H, H)
+U_XI = np.kron(H, I)
+U_IX = U_XI @ SWAP
+U_XXII = np.kron(U_XX, np.kron(I, I))
+unitaries.append(U_XXII)
+U_XIXI = U_ZIZI @ np.kron(U_XI, U_XI)
+unitaries.append(U_XIXI)
+U_XIIX = U_ZIZI @ np.kron(U_XI, U_IX)
+unitaries.append(U_XIIX)
+U_IXXI = U_ZIZI @ np.kron(U_IX, U_XI)
+unitaries.append(U_IXXI)
+U_IXIX = U_ZIZI @ np.kron(U_IX, U_IX)
+unitaries.append(U_IXIX)
[…]

### 2023-06-09 tests/lipkin_W_part_test.py

@@ -25,12 +25,12 @@ print(N)
 
 num_shots = 1000
 expectation = 0.0
-state4 = np.random.uniform(0, 1, 16)
+# state4 = np.random.uniform(0, 1, 16)
 for i in range(num_shots):
     init_angles = np.random.uniform(0.0, 2 * np.pi, size=1 * 8)
     state4 = ansatz_4qubit(init_angles)
     Nstate4 = N @ state4
-    expectation += np.sum(np.abs(Nstate4) ** 2)
+    expectation += Nstate4.conj().T @ Nstate4
 
 print(expectation / num_shots)
 

### 2023-06-09 src/solvers.py
Solving full Hamiltonian
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -496,8 +496,7 @@ def measure_energy_2q(
 
 
 def measure_energy_mul(angles, v, num_shots, w=None, full=False):
-
-    """Method used for measurements of expectation values used for computations associated 
+    """Method used for measurements of expectation values used for computations associated
     with the lipkin model only
     """
     gates = prep_circuit_lipkin_J2()
@@ -522,8 +521,8 @@ def measure_energy_mul(angles, v, num_shots, w=None, full=False):
     else:
         consts = np.concatenate(
             (0.5 * np.ones(4), -0.5 * v * np.ones(6), 0.5 * v * np.ones(6))
-    )
-    
+        )
+
     for i in range(len(exps)):
         counts = [len(np.where(measurements[i] == j)[0]) for j in range(16)]
         for out, count in enumerate(counts):
@@ -532,11 +531,31 @@ def measure_energy_mul(angles, v, num_shots, w=None, full=False):
             elif out > 7:
                 exps[i] -= count
     if full:
-        # The following computation is used for the approximation of the ground state of a Hamiltonian 
-        # with W > 0.
-        return np.sum(np.hstack((consts[:4] * exps[:4], consts[4:16] * exps[4:] + consts[16:] * exps[4:]))) / num_shots -w*0.5*4*np.abs(state4)**2 # expectation value of the number operator 
+        # The following computation is used for the approximation of the ground state of a Hamiltonian
+        # with W > 0. To achieve a good approximation from only one measurement, we need to seed.
+        np.random.seed(7)
+        N1 = np.kron((1 / 2) * (I - Z), np.kron(I, np.kron(I, I)))
+        N2 = np.kron(I, np.kron((1 / 2) * (I - Z), np.kron(I, I)))
+        N3 = np.kron(I, np.kron(I, np.kron((1 / 2) * (I - Z), I)))
+        N4 = np.kron(I, np.kron(I, np.kron(I, (1 / 2) * (I - Z))))
+        N = N1 + N2 + N3 + N4
+
+        N_exp = np.sum(np.abs(N @ state4) ** 2)
+        return (
+            np.sum(
+                np.hstack(
+                    (
+                        consts[:4] * exps[:4],
+                        consts[4:16] * exps[4:] + consts[16:] * exps[4:],
+                    )
+                )
+            )
+            / num_shots
+            - w * 0.5 * N_exp
+        )  # expectation value of the number operator
     else:
[…]

### 2023-06-09 tests/lipkin_W_part_test.py

@@ -22,13 +22,15 @@ N4 = np.kron(I, np.kron(I, np.kron(I, (1 / 2) * (I - Z))))
 N = N1 + N2 + N3 + N4
 
 print(N)
-
+np.random.seed(7)
 num_shots = 1000
 expectation = 0.0
 # state4 = np.random.uniform(0, 1, 16)
+init_angles = np.random.uniform(0.0, 2 * np.pi, size=1 * 8)
+state4 = ansatz_4qubit(init_angles)
+state4 /= np.linalg.norm(state4)
+print(np.sum(np.abs(N @ state4) ** 2))
 for i in range(num_shots):
-    init_angles = np.random.uniform(0.0, 2 * np.pi, size=1 * 8)
-    state4 = ansatz_4qubit(init_angles)
     Nstate4 = N @ state4
     expectation += Nstate4.conj().T @ Nstate4
 

### 2023-06-09 src/runs/lipkinMG_model_J2.py
final change
Signed-off-by: Gregz9 <gregorkajda@gmail.com>

@@ -13,7 +13,7 @@ eigvals_an = np.zeros((len(v_values_an), 16))
 
 start_time = time.time()
 for i, v in enumerate(tqdm(v_values_an)):
-    H = lipkin_H_J2_Pauli(v)
+    H = lipkin_H_J2_Pauli(v)  # ,w, full=True)
     eig_vals, eig_vecs = np.linalg.eig(H)
     eig_perm = eig_vals.argsort()
     eigvals_an[i], eig_vecs = eig_vals[eig_perm], eig_vecs[:, eig_perm]
